### 2024-10-27 paper/paper.qmd
pollster analysis

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)

data <- read.csv("../data/01-raw_data/raw_data.csv")
```

```{r}
# pollsters measured by highest numeric_grade, transparency_score, pollster_rating_id

usdata <- data %>%
  group_by(pollster) %>%
  summarize(
    pollster_rating_id = mean(pollster_rating_id, na.rm = TRUE),
    numeric_grade = mean(numeric_grade, na.rm = TRUE),
    pollscore = mean(pollscore, na.rm = TRUE),
    transparency_score = mean(transparency_score, na.rm = TRUE)
  ) %>%
  filter(numeric_grade >= 1.5, transparency_score >= 5, pollster_rating_id >= 400) %>%
    arrange(desc(pollster_rating_id))
usdata

# pollscore is negative ? how is it measured 
```
# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-10-30 paper/paper.qmd
test of model

@@
 library(dplyr)
-
-data <- read.csv("../data/01-raw_data/raw_data.csv")
+library(janitor)
 ```
@@
 ```{r}
-# pollsters measured by highest numeric_grade, transparency_score, pollster_rating_id
+#| include: false
+#| warning: false
+#| message: false
 
-usdata <- data %>%
-  group_by(pollster) %>%
-  summarize(
-    pollster_rating_id = mean(pollster_rating_id, na.rm = TRUE),
-    numeric_grade = mean(numeric_grade, na.rm = TRUE),
-    pollscore = mean(pollscore, na.rm = TRUE),
-    transparency_score = mean(transparency_score, na.rm = TRUE)
+#Load and clean data
+data <- read.csv("../data/01-raw_data/raw_data.csv") |>
+  clean_names()
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter NYT polls and remove entries with NA or empty value for statr
+# Select imporntat variables (change as you please)
+
+preddata <- data %>%
+ # filter(pollster_id == 1424 & !is.na(state) & state != "") %>%
+  select(display_name, question_id, sample_size, state, party, answer, candidate_id, candidate_name, pct, start_date, end_date)
+
+head(preddata)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+#FILTERED SWING STATES, SUBJECT TO CHANGE, ALSO INCLUDED NATIONAL?
+
+harrisdata <- preddata %>%
+  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "")) %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  mutate(
+    state = if_else(state == "", "National", state),
+    end_date = mdy(end_date)
   ) %>%
-  filter(numeric_grade >= 1.5, transparency_score >= 5, pollster_rating_id >= 400) %>%
-    arrange(desc(pollster_rating_id))
-usdata
 
-# pollscore is negative ? how is it measured 
+  # Step 4: Filter by date and calculate number of respondents
+  filter(end_date >= as.Date("2024-07-21")) %>%
+  mutate(
+    num_harris = round((pct / 100) * sample_size, 0)
+  )
+
+# Plot the filtered and transformed data
+base_plot <- ggplot(harrisdata, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date") +
+  geom_point(aes(color = state)) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+base_plot
+
+base_plot +
+  geom_point() +
+  geom_smooth()
+
 ```
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+## MODEL 1, LINEAR MODEL WITH PREDICTOR END_DATE
+
+model_date <- lm(pct ~ end_date, data = harrisdata)
+
+harrisdata <- harrisdata |>
+  mutate(
+    fitted_date = predict(model_date)
+  )
+
+ggplot(harrisdata, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
+  geom_line(aes(y = fitted_date), color = "blue", linetype = "dotted") +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date")
+```
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+## MODEL 2, BAYESIAN MODEL WITH PREDICTOR END_DATE
+
+harrisdata <- harrisdata |>
+  mutate(state = factor(state))
+
+model1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | state)
+
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model_1 <- stan_glmer(
+  formula = model1,
+  data = harrisdata,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+pp_check(bayesian_model_1)
+
+summary(bayesian_model_1)
+
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+
+```
+
+```{r}
+harrisdata <- harrisdata |>
+  mutate(
+    end_date_num = as.numeric(end_date - min(end_date))
+  )
+
+spline_model <- stan_glm(
+  pct ~ ns(end_date_num, df = 6),
+  data = harrisdata,
+  family = gaussian(),
+  prior = normal(0, 5),
+  prior_intercept = normal(50, 10),
+  seed = 1234,
+  iter = 2000,
+  chains = 4,
+  refresh = 0
+)
+
+summary(spline_model)
+pp_check(spline_model)
+
+# Summarize the model
+summary(spline_model)
+
+# Posterior predictive checks
+pp_check(spline_model)
+```
+
+```{r}
+new_data <- data.frame(
+  end_date_num = seq(
+    min(harrisdata$end_date_num),
+    max(harrisdata$end_date_num),
+    length.out = 100))
+
+posterior_preds <- posterior_predict(spline_model, newdata = new_data)
+
+pred_summary <- new_data |>
+  mutate(
+    pred_mean = colMeans(posterior_preds),
+    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
+    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
+  )
+
+ggplot(harrisdata, aes(x = end_date_num, y = pct)) +
+  geom_point() +
+  geom_line(
+    data = pred_summary,
+    aes(x = end_date_num, y = pred_mean),
+    color = "blue",
+    inherit.aes = FALSE
+  ) +
+  geom_ribbon(
+    data = pred_summary,
+    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
+    alpha = 0.2,
+    inherit.aes = FALSE
+  ) +
+  labs(
+    x = "Days since earliest poll",
+    y = "Percentage",
+    title = "Poll Percentage over Time with Spline Fit"
+  ) +
+  theme_minimal()
+
+```
+
 # Introduction
@@
 
+Why we choose NYT?
+- `numeric_grade` is 3
+- `pollscore` is -1.5, a score of reliability called "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.", where negative numbers are better
+- `transparency_score` is 9, reflects pollsters transparency about their methodology (calculated based on how much information it discloses about its polls and weighted by recency)
+- `population_full` is 'rv', respondents are registered voters
+
 In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

### 2024-10-31 paper/paper.qmd
MODEL UPDATE!

--- 2024-10-27 paper/paper.qmd
@@
 library(dplyr)
-
-data <- read.csv("../data/01-raw_data/raw_data.csv")
+library(janitor)
+library(lubridate)
+library(broom)
+library(modelsummary)
+library(rstanarm)
+library(splines)
 ```
@@
 ```{r}
-# pollsters measured by highest numeric_grade, transparency_score, pollster_rating_id
+#| include: false
+#| warning: false
+#| message: false
 
-usdata <- data %>%
-  group_by(pollster) %>%
-  summarize(
-    pollster_rating_id = mean(pollster_rating_id, na.rm = TRUE),
-    numeric_grade = mean(numeric_grade, na.rm = TRUE),
-    pollscore = mean(pollscore, na.rm = TRUE),
-    transparency_score = mean(transparency_score, na.rm = TRUE)
+#Load and clean data
+data <- read.csv("../data/01-raw_data/raw_data.csv") |>
+  clean_names()
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter NYT polls and remove entries with NA or empty value for statr
+# Select imporntat variables (change as you please)
+
+#& !is.na(state) & state != ""
+
+preddata <- data %>%
+  filter(pollster_id == 1424) %>%
+  select(display_name, question_id, sample_size, state, party, answer, candidate_id, candidate_name, pct, start_date, end_date)
+
+print(preddata)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+#FILTERED SWING STATES, SUBJECT TO CHANGE, ALSO INCLUDED NATIONAL?
+harrisdata <- preddata %>%
+  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  mutate(
+    end_date = mdy(end_date)    #is date mutated correctly here 
   ) %>%
-  filter(numeric_grade >= 1.5, transparency_score >= 5, pollster_rating_id >= 400) %>%
-    arrange(desc(pollster_rating_id))
-usdata
 
-# pollscore is negative ? how is it measured 
+  # Step 4: Filter by date and calculate number of respondents
+  filter(end_date >= as.Date("2024-07-21")) %>%
+  mutate(
+    num_harris = round((pct / 100) * sample_size, 0)
+  )
+
+harrisdata
+
+#### Plot data ####
+base_plot <- ggplot(harrisdata, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date") +
+    theme(legend.position = "bottom")
+
+# Plots poll estimates and overall smoothing
+base_plot +
+  geom_point() +
+  geom_smooth()
+
+# This gets messy - need to add a filter - see line 21
+base_plot +
+  geom_point(aes(color = state)) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
 ```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+## MODEL 1, BAYESIAN MODEL WITH PREDICTOR STATE
+
+# Define Model
+model1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | state)
+
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model1 <- stan_glmer(
+  formula = model1,
+  data = harrisdata,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Plot checks for model 1
+pp_check(bayesian_model1)
+summary(bayesian_model1)
+plot(bayesian_model1, pars = "(Intercept)", prob = 0.95)
+print(modelsummary(bayesian_model1))
+```
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+## MODEL 2, BAYESIAN MODEL WITH PREDICTOR END_DATE AND STATE
+
+# Turn end_date into quantifiable column, to use for model
+harrisdata <- harrisdata %>%
+  mutate(
+    end_date_num = as.numeric(end_date - min(end_date)))
+
+harrisdata <- harrisdata |>
+  mutate(state = factor(state))
+
+# Define Model 2
+model2 <- cbind(num_harris, sample_size - num_harris) ~ (1 | end_date_num) + (1 | state)
+
+bayesian_model2 <- stan_glmer(
+  formula = model2,
+  data = harrisdata,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Plot checks for Model 2
+pp_check(bayesian_model2)
+summary(bayesian_model2)
+plot(bayesian_model2, pars = "(Intercept)", prob = 0.95)
+print(modelsummary(bayesian_model2))
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Predict and plot
+
+spline_model_pred <- stan_glm(
+  pct ~ ns(end_date_num, df = 5) + state,
+  data = harrisdata,
+  family = gaussian(),
+  prior = normal(0, 5),
+  prior_intercept = normal(50, 10),
+  seed = 1234,
+  iter = 2000,
+  chains = 4,
+  refresh = 0
+)
+
+# Summarize the model
+summary(spline_model_pred)
+
+# Posterior predictive checks
+pp_check(spline_model_pred)
+```
+
+```{r}
+
+new_data <- data.frame(
+  end_date_num = seq(
+    min(harrisdata$end_date_num),
+    max(harrisdata$end_date_num),
+    length.out = 100),
+  state = factor("Georgia", levels = levels(harrisdata$state)))
+
+posterior_preds <- posterior_predict(spline_model_pred, newdata = new_data)
+
+pred_summary <- new_data |>
+  mutate(
+    pred_mean = colMeans(posterior_preds),
+    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
+    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975))
+
+ggplot(harrisdata, aes(x = end_date_num, y = pct, color = state)) +
+  geom_point() +
+  geom_line(
+    data = pred_summary,
+    aes(x = end_date_num, y = pred_mean),
+    color = "blue",
+    inherit.aes = FALSE) +
+  geom_ribbon(
+    data = pred_summary,
+    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
+    alpha = 0.2,
+    inherit.aes = FALSE) +
+  labs(
+    x = "Days since earliest poll",
+    y = "Percentage",
+    title = "Poll Percentage over Time with Spline Fit") +
+  theme_minimal()
+```
+
 # Introduction

### 2024-11-01 paper/paper.qmd
commit tp ppull

--- 2024-10-27 paper/paper.qmd
@@
   - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+thanks: "Code and data are available at: [https://github.com/Mezhi18/US_Election2024.git](https://github.com/Mezhi18/US_Election2024.git)."
 date: today

### 2024-11-02 paper/paper.qmd
model ideas
Wrote about data

--- 2024-10-31 paper/paper.qmd
@@
 ```{r}
-#| echo: false
+#| include: false
 #| warning: false
@@
 
+As our paper is about the 2024 United States federal election and more specifically we are looking at the pollsand the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election.
+Our orignal Data set had over 16,000 unique entries from different pollsters, the business or Organization that conducts the poll.
+Each poll has two entries one giving the Data for the polling a opinions of Donald Trump and the second for Kamala Harris.
+
+As there are over 50 variables many of which are redundent to our paper we will only discuss those that we have kept in our clean data as they are the only ones we use in our analysis.
+
+**display_name:** This column shows the name of the Pollster that conducted the poll, as we want to keep consistency we are only looking at polls conducted by the New York Times.
+
+**question_id:** This varible refers to the question that was asked in the specific poll.
+
+**sample_size:** This variable is the quantity of people that participated in the specific poll.
+
 ## Overview
@@
 Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
 

### 2024-11-02 paper/paper.qmd
Additional info about data

@@
 
-As our paper is about the 2024 United States federal election and more specifically we are looking at the pollsand the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election.
+As our paper is about the 2024 United States federal election and more specifically we are looking at the polls and the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election.
 Our orignal Data set had over 16,000 unique entries from different pollsters, the business or Organization that conducts the poll.
 Each poll has two entries one giving the Data for the polling a opinions of Donald Trump and the second for Kamala Harris.
+We have aquired our polling data from @bibnotmadeyet.
 
@@
 **sample_size:** This variable is the quantity of people that participated in the specific poll.
+
+**state:** This variable tells us in which States the poll was conducted.
+
+**party:** This variable tells us which party the poll participants intend to vot for.
+
+**answer:** This variable tells us the candidate the participant intend to vote for.
+
+**candidate_id:** This variable is the numerical value assigned to each candidate.
+
+**candidate_name**: This is the full name of the selected candidate.
+
+**pct:** This tells us the percentage of participents that intend to vote for the selected candidate.
+
+**start_date:** The date the pollster began conducting the poll.
+
+**end_date:** The date the pollster finished conducting the poll.
+
+**num_harris:** The numbner of participants that intend to vot for Kamala Harris.
+
+**end_date_num:** IDK
 

### 2024-11-02 paper/paper.qmd
Started introduction

@@
 
-Overview paragraph
+The 2024 United States presidential election represents a pivotal moment in the country's political landscape.
+As in previous elections, swing states are projected to play a critical role in determining the outcome.
+Swing states, which are characterized by their shifting voting patterns and balanced support for both major political parties, have historically been the focus of intense campaign efforts and polling analyses.
+Understanding the dynamics and voter preferences in these states is crucial to gaining insight into the broader electoral trends that could shape the nation's future.
+
+This paper aims to analyze polling data related to the 2024 election, with a particular emphasis on the swing states.
+By examining various polls and identifying patterns in voter sentiment, we seek to uncover the factors that may influence voter behavior in these highly contested regions.
+Our analysis will explore demographic shifts, the impact of key issues, and the level of voter engagement across different swing states.
+Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the forces shaping the 2024 presidential election.
 
@@
 As our paper is about the 2024 United States federal election and more specifically we are looking at the polls and the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election.
-Our orignal Data set had over 16,000 unique entries from different pollsters, the business or Organization that conducts the poll.
+Our original Data set had over 16,000 unique entries from different pollsters, the business or Organization that conducts the poll.
 Each poll has two entries one giving the Data for the polling a opinions of Donald Trump and the second for Kamala Harris.
-We have aquired our polling data from @bibnotmadeyet.
+We have acquired our polling data from @bibnotmadeyet.
 
-As there are over 50 variables many of which are redundent to our paper we will only discuss those that we have kept in our clean data as they are the only ones we use in our analysis.
+As there are over 50 variables many of which are redundant to our paper we will only discuss those that we have kept in our clean data as they are the only ones we use in our analysis.
 
@@
 
-**question_id:** This varible refers to the question that was asked in the specific poll.
+**question_id:** This variable refers to the question that was asked in the specific poll.
 
@@
 
-**pct:** This tells us the percentage of participents that intend to vote for the selected candidate.
+**pct:** This tells us the percentage of participants that intend to vote for the selected candidate.
 
@@
 
-**num_harris:** The numbner of participants that intend to vot for Kamala Harris.
+**num_harris:** The number of participants that intend to vote for Kamala Harris.
 
@@
 
+```{=tex}
 \begin{align} 
@@
 \end{align}
-
+```
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
@@
 Why we choose NYT?
-- `numeric_grade` is 3
-- `pollscore` is -1.5, a score of reliability called "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.", where negative numbers are better
-- `transparency_score` is 9, reflects pollsters transparency about their methodology (calculated based on how much information it discloses about its polls and weighted by recency)
-- `population_full` is 'rv', respondents are registered voters
+- `numeric_grade` is 3 - `pollscore` is -1.5, a score of reliability called "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.", where negative numbers are better - `transparency_score` is 9, reflects pollsters transparency about their methodology (calculated based on how much information it discloses about its polls and weighted by recency) - `population_full` is 'rv', respondents are registered voters
 

### 2024-11-02 paper/paper.qmd
Discussion Topics: Swing states
test
model and graphs

@@
 
-# Filter NYT polls and remove entries with NA or empty value for statr
 # Select imporntat variables (change as you please)
 
-#& !is.na(state) & state != ""
-
 preddata <- data %>%
-  filter(pollster_id == 1424) %>%
-  select(display_name, question_id, sample_size, state, party, answer, candidate_id, candidate_name, pct, start_date, end_date)
+  filter(numeric_grade >1) %>%
+  select(pollster, sample_size, state, party, answer, candidate_id, candidate_name, pct, start_date, end_date, population, pollscore, numeric_grade) %>%
+  filter(!is.na(pollscore)) %>%
+  filter(!is.na(state) & state != "")
 
 print(preddata)
+
 ```
@@
   ) %>%
-
-  # Step 4: Filter by date and calculate number of respondents
   filter(end_date >= as.Date("2024-07-21")) %>%
@@
 
-## MODEL 1, BAYESIAN MODEL WITH PREDICTOR STATE
-
-# Define Model
-model1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | state)
-
-priors <- normal(0, 2.5, autoscale = TRUE)
-
-bayesian_model1 <- stan_glmer(
-  formula = model1,
-  data = harrisdata,
-  family = binomial(link = "logit"),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95
-)
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-# Plot checks for model 1
-pp_check(bayesian_model1)
-summary(bayesian_model1)
-plot(bayesian_model1, pars = "(Intercept)", prob = 0.95)
-print(modelsummary(bayesian_model1))
-```
-
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
 ## MODEL 2, BAYESIAN MODEL WITH PREDICTOR END_DATE AND STATE
@@
 
-# Define Model 2
-model2 <- cbind(num_harris, sample_size - num_harris) ~ (1 | end_date_num) + (1 | state)
+# Define Model 1
+model <- cbind(num_harris, sample_size - num_harris) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
 
-bayesian_model2 <- stan_glmer(
-  formula = model2,
+bayesian_model <- stan_glmer(
+  formula = model,
   data = harrisdata,
@@
   cores = 4,
-  adapt_delta = 0.95
-)
+  adapt_delta = 0.95)
 ```
@@
 
-# Plot checks for Model 2
-pp_check(bayesian_model2)
-summary(bayesian_model2)
-plot(bayesian_model2, pars = "(Intercept)", prob = 0.95)
-print(modelsummary(bayesian_model2))
+# Plot checks for Model
+pp_check(bayesian_model)
+summary(bayesian_model)
+plot(bayesian_model, pars = "(Intercept)", prob = 0.95)
+print(modelsummary(bayesian_model))
+summary(bayesian_model)
 ```
@@
 spline_model_pred <- stan_glm(
-  pct ~ ns(end_date_num, df = 5) + state,
+  pct ~ ns(end_date_num, df = 5) + state + pollster,
   data = harrisdata,
@@
 ```{r}
+#| echo: false
+#| warning: false
+#| message: false
 
+# Convert `pollster` to a factor with defined levels if it's not already
+harrisdata$pollster <- factor(harrisdata$pollster, levels = unique(harrisdata$pollster))
+
+# Define the new data grid for prediction, ensuring each factor has the correct levels
 new_data <- data.frame(
@@
     max(harrisdata$end_date_num),
-    length.out = 100),
+    length.out = 100
+  ),
+   pollster = factor("YouGov", levels = levels(harrisdata$pollster)),
   state = factor("Georgia", levels = levels(harrisdata$state)))
@@
     pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
-    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975))
+    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
+  )
 
@@
     title = "Poll Percentage over Time with Spline Fit") +
+   facet_wrap(~ state, scales = "free_y") +
   theme_minimal()
@@
 
-This paper aims to analyze polling data related to the 2024 election, with a particular emphasis on the swing states.
+In this paper we aim to analyze polling data related to the 2024 election, with a particular emphasis on the swing states.
 By examining various polls and identifying patterns in voter sentiment, we seek to uncover the factors that may influence voter behavior in these highly contested regions.
@@
 
-## First discussion point {#sec-first-point}
+## Arizona {#sec-arizona}
 
@@
 
-## Second discussion point
+## Georgia {#sec-georgia}
 
@@
 
-## Third discussion point
+## Michigan {#sec-michigan}
+
+## Nevada {#sec-nevada}
+
+## North Carolina {#sec-norcar}
+
+## Michigan {#sec-michigan}
+
+## Pennsylvania {#sec-penn}
+
+## Wisconsin {#sec-Wisconsin}
 

### 2024-11-02 paper/paper.qmd
commiting to pull

--- 2024-11-02 paper/paper.qmd
@@
 ```{r}
-#| include: false
+#| echo: false
 #| warning: false
@@
 
-As our paper is about the 2024 United States federal election and more specifically we are looking at the polls and the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election.
-Our orignal Data set had over 16,000 unique entries from different pollsters, the business or Organization that conducts the poll.
-Each poll has two entries one giving the Data for the polling a opinions of Donald Trump and the second for Kamala Harris.
-We have aquired our polling data from @bibnotmadeyet.
-
-As there are over 50 variables many of which are redundent to our paper we will only discuss those that we have kept in our clean data as they are the only ones we use in our analysis.
-
-**display_name:** This column shows the name of the Pollster that conducted the poll, as we want to keep consistency we are only looking at polls conducted by the New York Times.
-
-**question_id:** This varible refers to the question that was asked in the specific poll.
-
-**sample_size:** This variable is the quantity of people that participated in the specific poll.
-
-**state:** This variable tells us in which States the poll was conducted.
-
-**party:** This variable tells us which party the poll participants intend to vot for.
-
-**answer:** This variable tells us the candidate the participant intend to vote for.
-
-**candidate_id:** This variable is the numerical value assigned to each candidate.
-
-**candidate_name**: This is the full name of the selected candidate.
-
-**pct:** This tells us the percentage of participents that intend to vote for the selected candidate.
-
-**start_date:** The date the pollster began conducting the poll.
-
-**end_date:** The date the pollster finished conducting the poll.
-
-**num_harris:** The numbner of participants that intend to vot for Kamala Harris.
-
-**end_date_num:** IDK
-
 ## Overview
@@
 Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
 
@@
 
+# Appendix 1
+
+# Appendix 2
+
+\newpage
+
 # References

### 2024-11-03 paper/paper.qmd
model update with vizualisation

--- 2024-11-02 paper/paper.qmd
@@
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
+format:
+  pdf:
+    bibliography: references.bib
+    documentclass: article
+    geometry: margin = 1in
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 library(splines)
-```
-
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-#Load and clean data
-data <- read.csv("../data/01-raw_data/raw_data.csv") |>
-  clean_names()
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-# Select imporntat variables (change as you please)
-
-preddata <- data %>%
-  filter(numeric_grade >1) %>%
-  select(pollster, sample_size, state, party, answer, candidate_id, candidate_name, pct, start_date, end_date, population, pollscore, numeric_grade) %>%
-  filter(!is.na(pollscore)) %>%
-  filter(!is.na(state) & state != "")
-
-print(preddata)
-
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-#FILTERED SWING STATES, SUBJECT TO CHANGE, ALSO INCLUDED NATIONAL?
-harrisdata <- preddata %>%
-  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
-  filter(candidate_name == "Kamala Harris") %>%
-  mutate(
-    end_date = mdy(end_date)    #is date mutated correctly here
-  ) %>%
-  filter(end_date >= as.Date("2024-07-21")) %>%
-  mutate(
-    num_harris = round((pct / 100) * sample_size, 0)
-  )
-
-harrisdata
-
-#### Plot data ####
-base_plot <- ggplot(harrisdata, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date") +
-    theme(legend.position = "bottom")
-
-# Plots poll estimates and overall smoothing
-base_plot +
-  geom_point() +
-  geom_smooth()
-
-# This gets messy - need to add a filter - see line 21
-base_plot +
-  geom_point(aes(color = state)) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
-
-```
-
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-## MODEL 2, BAYESIAN MODEL WITH PREDICTOR END_DATE AND STATE
-
-# Turn end_date into quantifiable column, to use for model
-harrisdata <- harrisdata %>%
-  mutate(
-    end_date_num = as.numeric(end_date - min(end_date)))
-
-harrisdata <- harrisdata |>
-  mutate(state = factor(state))
-
-# Define Model 1
-model <- cbind(num_harris, sample_size - num_harris) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
-
-bayesian_model <- stan_glmer(
-  formula = model,
-  data = harrisdata,
-  family = binomial(link = "logit"),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-# Plot checks for Model
-pp_check(bayesian_model)
-summary(bayesian_model)
-plot(bayesian_model, pars = "(Intercept)", prob = 0.95)
-print(modelsummary(bayesian_model))
-summary(bayesian_model)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-# Predict and plot
-
-spline_model_pred <- stan_glm(
-  pct ~ ns(end_date_num, df = 5) + state + pollster,
-  data = harrisdata,
-  family = gaussian(),
-  prior = normal(0, 5),
-  prior_intercept = normal(50, 10),
-  seed = 1234,
-  iter = 2000,
-  chains = 4,
-  refresh = 0
-)
-
-# Summarize the model
-summary(spline_model_pred)
-
-# Posterior predictive checks
-pp_check(spline_model_pred)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-# Convert `pollster` to a factor with defined levels if it's not already
-harrisdata$pollster <- factor(harrisdata$pollster, levels = unique(harrisdata$pollster))
-
-# Define the new data grid for prediction, ensuring each factor has the correct levels
-new_data <- data.frame(
-  end_date_num = seq(
-    min(harrisdata$end_date_num),
-    max(harrisdata$end_date_num),
-    length.out = 100
-  ),
-   pollster = factor("YouGov", levels = levels(harrisdata$pollster)),
-  state = factor("Georgia", levels = levels(harrisdata$state)))
-
-posterior_preds <- posterior_predict(spline_model_pred, newdata = new_data)
-
-pred_summary <- new_data |>
-  mutate(
-    pred_mean = colMeans(posterior_preds),
-    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
-    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
-  )
-
-ggplot(harrisdata, aes(x = end_date_num, y = pct, color = state)) +
-  geom_point() +
-  geom_line(
-    data = pred_summary,
-    aes(x = end_date_num, y = pred_mean),
-    color = "blue",
-    inherit.aes = FALSE) +
-  geom_ribbon(
-    data = pred_summary,
-    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
-    alpha = 0.2,
-    inherit.aes = FALSE) +
-  labs(
-    x = "Days since earliest poll",
-    y = "Percentage",
-    title = "Poll Percentage over Time with Spline Fit") +
-   facet_wrap(~ state, scales = "free_y") +
-  theme_minimal()
 ```
@@
 
+## Overview
+
 As our paper is about the 2024 United States federal election and more specifically we are looking at the polls and the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election.
 Our original Data set had over 16,000 unique entries from different pollsters, the business or Organization that conducts the poll.
-Each poll has two entries one giving the Data for the polling a opinions of Donald Trump and the second for Kamala Harris.
+Each poll has two entries, one giving the Data for the polling opinions of Donald Trump and the second for Kamala Harris.
 We have acquired our polling data from @bibnotmadeyet.
@@
 
-**display_name:** This column shows the name of the Pollster that conducted the poll, as we want to keep consistency we are only looking at polls conducted by the New York Times.
+- `pollster`: Shows name of the Pollster that conducted the poll.
+- `sample_size`: The number of people that participated in the specific poll.
+- `state`: This variable tells us in which States the poll was conducted.
+- `candidate_name`: This is the full name of the selected candidate.
+- `pct`: This tells us the percentage of participants that intend to vote for the selected candidate.
+- `start_date`: The date the pollster began conducting the poll.
+- `end_date`: The date the pollster finished conducting the poll.
 
-**question_id:** This variable refers to the question that was asked in the specific poll.
+The variables that we have create are:
 
-**sample_size:** This variable is the quantity of people that participated in the specific poll.
+- `num_harris`: The number of participants that intend to vote for Kamala Harris.
+- `end_date_num`: The number of days since the first poll since Harris announced her candidacy.
 
-**state:** This variable tells us in which States the poll was conducted.
+Each pollster has a numeric grade from 1.0 to 3.0, which indicates the quality/ reliability of the respective pollster.
+Additionally, each pollster is also given a transparency score from 1.0 to 10.0 reflecting how 'transparent' the pollster is, or how much information is disclosed about its polls and methodology.
+It is important for pollsters to maintain high numeric grades and transparency scores because these metrics directly reflect the quality and reliability of their data.
+To ensure the highest level of accuracy in our predictions, we only include polls with a numeric grade of 1.5 or above and a transparency score of 6.0 or above.
 
-**party:** This variable tells us which party the poll participants intend to vot for.
-
-**answer:** This variable tells us the candidate the participant intend to vote for.
-
-**candidate_id:** This variable is the numerical value assigned to each candidate.
-
-**candidate_name**: This is the full name of the selected candidate.
-
-**pct:** This tells us the percentage of participants that intend to vote for the selected candidate.
-
-**start_date:** The date the pollster began conducting the poll.
-
-**end_date:** The date the pollster finished conducting the poll.
-
-**num_harris:** The number of participants that intend to vote for Kamala Harris.
-
-**end_date_num:** IDK
-
-## Overview
+¿ does this go into data cleaning ?
 
@@
 
-Overview text
+## Methodology
 
-## Measurement
+This paper uses 'polls-of-polls' method to analyze and predict our outcome, which combines polls from multiple sources (pollsters).
+By combining results from various sources, this approach incorporates diverse perspectives, which helps to minimize individual biases and provides a more balanced view.
+Unlike relying on a single poll, which can be influenced by its own biases, the 'polls-of-polls' method enhances reliability and increases the overall validity of the results.
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+... explain how pollsters survey people, what their different methodologies are, and why these things are important for an accurate prediction...
+
+## Measurement/ Data Visualization
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+#Load and clean data
+data <- read.csv("../data/01-raw_data/raw_data.csv") |>
+  clean_names()
+
+preddata <- data %>%
+  filter(numeric_grade >1.5 & transparency_score >6) %>%
+  select(pollster, sample_size, state, candidate_name, pct, start_date, end_date, pollscore, numeric_grade) %>%
+  filter(!is.na(pollscore)) %>%
+  filter(!is.na(state) & state != "")
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-supportstate
+#| fig-cap: "Distribution of Support Percentage by State"
+
+harrisdata <- preddata %>%
+  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  mutate(end_date = mdy(end_date)) %>%
+  filter(end_date >= as.Date("2024-07-21")) %>%
+  mutate(num_harris = round((pct / 100) * sample_size, 0))
+
+ggplot(harrisdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
+  geom_boxplot() +
+  coord_flip() +  # Flip for better readability
+  labs(x = "State", y = "Support Percentage") +
+  theme_minimal() +
+  theme(legend.position = "none")
+```
+We focus on the swing states of the U.S, according to  [cite NYT PAPER]
+
+These are the statistics of each swing state by percetnage by state.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Plot data ####
+base_plot <- ggplot(harrisdata, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date") +
+    theme(legend.position = "bottom")
+
+base_plot +
+  geom_point(aes(color = state)) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+ggplot(harrisdata, aes(x = end_date, y = pct, color = state)) +
+  geom_point() +
+  geom_smooth(se = TRUE) +
+  facet_wrap(~ state, scales = "free_y") +
+  labs(y = "Harris Percent", x = "Date", title = "Polling Trends for Kamala Harris by State") +
+  theme_minimal()
+
+```
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
 Talk more about it.
@@
 (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
 
@@
 
+This is Model 1, follows Bayesian Model
+
+We define our model as:
+
+*define each vairable here*
+```{=tex}
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \beta_0 + \beta_1 \times \mbox{State}_i + \beta_2 \times \mbox{Days Since First Poll}_i + \beta_3 \times \mbox{Pollster}_i \\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Exponential}(1)
+\end{align}
+```
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Turn end_date into quantifiable column, to use for model
+harrisdata <- harrisdata %>%
+  mutate(
+    end_date_num = as.numeric(end_date - min(end_date)))
+
+harrisdata <- harrisdata |>
+  mutate(state = factor(state))
+
+# Define Model
+model <- cbind(num_harris, sample_size - num_harris) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
+
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model <- stan_glmer(
+  formula = model,
+  data = harrisdata,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  weights = harrisdata$numeric_grade,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Plot checks for Model
+pp_check(bayesian_model)
+plot(bayesian_model, pars = "(Intercept)", prob = 0.95)
+modelsummary::modelsummary(
+  list(
+    "First model" = bayesian_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
+
+This is Model 2
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+spline_model <- stan_glm(
+  pct ~ ns(end_date_num, df = 5) + pollster + state, 
+  data = harrisdata,
+  family = gaussian(),
+  prior = normal(0, 5),
+  prior_intercept = normal(50, 10),
+  seed = 1234,
+  iter = 2000,
+  chains = 4,
+  refresh = 0
+)
+
+# Summarize the model
+summary(spline_model)
+
+# Posterior predictive checks
+pp_check(spline_model)
+
+modelsummary::modelsummary(
+  list(
+    "First model" = spline_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+harrisdata$pollster <- factor(harrisdata$pollster)
+harrisdata$state <- factor(harrisdata$state)
+
+new_data <- data.frame(
+  end_date_num = seq(
+    min(harrisdata$end_date_num),
+    max(harrisdata$end_date_num),
+    length.out = 100
+  ),
+  pollster = factor("YouGov", levels = levels(harrisdata$pollster)),
+  state = factor("Georgia", levels = levels(harrisdata$state)))
+
+posterior_preds <- posterior_predict(spline_model, newdata = new_data)
+
+pred_summary <- new_data |>
+  mutate(
+    pred_mean = colMeans(posterior_preds),
+    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
+    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
+  )
+
+# Plot data and Bayesian model predictions with confidence ribbon
+ggplot(harrisdata, aes(x = end_date_num, y = pct, color = state)) +
+  geom_point() +
+  geom_line(
+    data = pred_summary,
+    aes(x = end_date_num, y = pred_mean),
+    color = "blue",
+    inherit.aes = FALSE
+  ) +
+  geom_ribbon(
+    data = pred_summary,
+    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
+    alpha = 0.2,
+    inherit.aes = FALSE
+  ) +
+  labs(
+    x = "Days since earliest poll",
+    y = "Percentage",
+    title = "Poll Percentage over Time"
+  ) +
+  facet_wrap(~ state, scales = "free_y") +
+  theme_minimal()
+```
+
+# TRUMP #
+
+```{r}
+trumpdata <- preddata %>%
+  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
+  filter(candidate_name == "Donald Trump") %>%
+  mutate(end_date = mdy(end_date)) %>%
+  filter(end_date >= as.Date("2024-07-21")) %>%
+  mutate(num_trump = round((pct / 100) * sample_size, 0))
+
+ggplot(trumpdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
+  geom_boxplot() +
+  coord_flip() +  # Flip for better readability
+  labs(x = "State", y = "Support Percentage") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+trumpdata <- trumpdata %>%
+  mutate(
+    end_date_num = as.numeric(end_date - min(end_date)))
+
+trumpdata <- trumpdata |>
+  mutate(state = factor(state))
+
+# Define Model
+model2 <- cbind(num_trump, sample_size - num_trump) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
+
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model2 <- stan_glmer(
+  formula = model2,
+  data = trumpdata,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  weights = trumpdata$numeric_grade,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95)
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Plot checks for Model
+pp_check(bayesian_model2)
+plot(bayesian_model2, pars = "(Intercept)", prob = 0.95)
+modelsummary::modelsummary(
+  list(
+    "First model" = bayesian_model2
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
+
 The goal of our modelling strategy is twofold. Firstly,...
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-```{=tex}
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-```
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+What do we expect...
+We expect a predictions in the high 40% based on our current data.
 
@@
 
+```{r}
+harrisdata <- harrisdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_harris))
+trumpdata <- trumpdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_trump))
+trumpdata
+```
+
+```{r}
+library(dplyr)
+library(ggplot2)
+library(rstanarm)
+
+harris_data <- harrisdata %>%
+  mutate(
+    candidate_name = "Kamala Harris",
+    support_ratio = num_harris  # Assuming harris_support_ratio is already calculated
+  ) %>%
+  select(end_date_num, candidate_name, support_ratio)
+
+# Prepare Trump data with a unified structure
+trump_data <- trumpdata %>%
+  mutate(
+    candidate_name = "Donald Trump",
+    support_ratio = num_trump  # Assuming trump_support_ratio is already calculated
+  ) %>%
+  select(end_date_num, candidate_name, support_ratio)
+
+# Combine both into a single data frame
+combined_data <- bind_rows(harris_data, trump_data)
+
+library(ggplot2)
+library(ggplot2)
+
+# Assuming `combined_data` contains columns `end_date_num`, `support_ratio`, and `candidate_name`
+
+ggplot(combined_data, aes(x = end_date_num, y = support_ratio, color = candidate_name, fill = candidate_name)) +
+  # Shaded area around each line for overlap visualization
+  geom_ribbon(data = combined_data %>% filter(candidate_name == "Kamala Harris"),
+              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
+              alpha = 0.3, fill = "red") +
+  geom_ribbon(data = combined_data %>% filter(candidate_name == "Donald Trump"),
+              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
+              alpha = 0.3, fill = "blue") +
+
+  # Line plot for each candidate
+  geom_line(aes(group = candidate_name), size = 1) +
+
+  scale_color_manual(values = c("Kamala Harris" = "red", "Donald Trump" = "blue")) +  # Set line colors
+  labs(
+    x = "Days Since Earliest Poll", 
+    y = "Support Ratio",
+    color = "Candidate",
+    title = "Support Ratio for Harris and Trump Over Time with Overlap Highlighted"
+  ) +
+  theme_minimal() +
+  theme(
+    text = element_text(size = 12),
+    plot.title = element_text(hjust = 0.5),
+    legend.position = "top",
+    legend.title = element_text(face = "bold")
+  )
+
+```
 # Discussion
@@
 ```{r}
-#| eval: true
-#| echo: false
+#| eval: false
 #| message: false
@@
 #| echo: false
-#| eval: true
+#| eval: false
 #| message: false
@@
 
+# Appendix 1
+
+# Appendix 2
+
+=======
+>>>>>>> eda40e1517683f0ad12e0bcd901797a2bb2abd7d
+\newpage
+
 # References

### 2024-11-03 paper/paper.qmd
added appenidx 2

@@
 ```
-We focus on the swing states of the U.S, according to  [cite NYT PAPER]
 
@@
 ```
-
-# TRUMP #
-
-```{r}
-trumpdata <- preddata %>%
-  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
-  filter(candidate_name == "Donald Trump") %>%
-  mutate(end_date = mdy(end_date)) %>%
-  filter(end_date >= as.Date("2024-07-21")) %>%
-  mutate(num_trump = round((pct / 100) * sample_size, 0))
-
-ggplot(trumpdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
-  geom_boxplot() +
-  coord_flip() +  # Flip for better readability
-  labs(x = "State", y = "Support Percentage") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-trumpdata <- trumpdata %>%
-  mutate(
-    end_date_num = as.numeric(end_date - min(end_date)))
-
-trumpdata <- trumpdata |>
-  mutate(state = factor(state))
-
-# Define Model
-model2 <- cbind(num_trump, sample_size - num_trump) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
-
-priors <- normal(0, 2.5, autoscale = TRUE)
-
-bayesian_model2 <- stan_glmer(
-  formula = model2,
-  data = trumpdata,
-  family = binomial(link = "logit"),
-  prior = priors,
-  prior_intercept = priors,
-  weights = trumpdata$numeric_grade,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-# Plot checks for Model
-pp_check(bayesian_model2)
-plot(bayesian_model2, pars = "(Intercept)", prob = 0.95)
-modelsummary::modelsummary(
-  list(
-    "First model" = bayesian_model2
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
 The goal of our modelling strategy is twofold. Firstly,...
@@
 
-```{r}
-harrisdata <- harrisdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_harris))
-trumpdata <- trumpdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_trump))
-trumpdata
-```
-
-```{r}
-library(dplyr)
-library(ggplot2)
-library(rstanarm)
-
-harris_data <- harrisdata %>%
-  mutate(
-    candidate_name = "Kamala Harris",
-    support_ratio = num_harris  # Assuming harris_support_ratio is already calculated
-  ) %>%
-  select(end_date_num, candidate_name, support_ratio)
-
-# Prepare Trump data with a unified structure
-trump_data <- trumpdata %>%
-  mutate(
-    candidate_name = "Donald Trump",
-    support_ratio = num_trump  # Assuming trump_support_ratio is already calculated
-  ) %>%
-  select(end_date_num, candidate_name, support_ratio)
-
-# Combine both into a single data frame
-combined_data <- bind_rows(harris_data, trump_data)
-
-library(ggplot2)
-library(ggplot2)
-
-# Assuming `combined_data` contains columns `end_date_num`, `support_ratio`, and `candidate_name`
-
-ggplot(combined_data, aes(x = end_date_num, y = support_ratio, color = candidate_name, fill = candidate_name)) +
-  # Shaded area around each line for overlap visualization
-  geom_ribbon(data = combined_data %>% filter(candidate_name == "Kamala Harris"),
-              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
-              alpha = 0.3, fill = "red") +
-  geom_ribbon(data = combined_data %>% filter(candidate_name == "Donald Trump"),
-              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
-              alpha = 0.3, fill = "blue") +
-
-  # Line plot for each candidate
-  geom_line(aes(group = candidate_name), size = 1) +
-
-  scale_color_manual(values = c("Kamala Harris" = "red", "Donald Trump" = "blue")) +  # Set line colors
-  labs(
-    x = "Days Since Earliest Poll", 
-    y = "Support Ratio",
-    color = "Candidate",
-    title = "Support Ratio for Harris and Trump Over Time with Overlap Highlighted"
-  ) +
-  theme_minimal() +
-  theme(
-    text = element_text(size = 12),
-    plot.title = element_text(hjust = 0.5),
-    legend.position = "top",
-    legend.title = element_text(face = "bold")
-  )
-
-```
 # Discussion
@@
 
-=======
->>>>>>> eda40e1517683f0ad12e0bcd901797a2bb2abd7d
+Here we will be talking about how we would conduct a $100,000 surevey to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent.
+Out of the \$100,000 we will allocate \$15000 for the necessary development and administration of the data.
+This means we will be spending this money to create the survey, and have a infrastructure in place to hold the large amount of data as well as cover any security fees to keep this data private and safe.
+Then we allocate $50000 to advertise the survey.
+We will need a large sample so we will be spending most of our budget for this.
+We will have a different urls for each advertisement we have so that we can incorporate this in our data to see which demographic is accessing the survey through which platform.
+The platforms to advertise this is through spotify, facebook, instagram, and some news networks.
+If possible we will try to use this money to get it endorsed by branches of government to show our reliability.
+Reliability is also what citizens respond to so this will get a lot of respondants.
+We will spend \$25,000 for modeling our data as they tend to be expensive with data this size.
+And the rest of the \$15,000 will be leftover cost for anything that we don't foresee.
+If there is any leftover we can add a survey participation price such as having a free 3 month trial for the platforms we mentioned above (that is if they have a subscription based membership).
+
+We will not be using telephone surveys because according to reserach [@idSurvey] we find the most people, don't pick up calls from unknown numbers and even the people that pick it up they are less inclined to answer the questions of the survey.
+
+Next to look at the actual contents of the survey.
+They can be accessed through this link:[Sample survey question](https://forms.office.com/Pages/ResponsePage.aspx?id=JsKqeAMvTUuQN7RtVsVSEL7G89G56QlHp_qqIPzGbJdUME9UQzJaWUlCVldZTktRV09UQ0Q3N0hQRi4u).
+There are 3 things that we were careful of when we were creating this survey.
+First thing we considered was transperency.
+People need the reassurence that the data that is being collected will not be used against them, and so we teel them what data we do collect and how the data we collect cannot be used to identify a person.
+The second thing we focus on is readability of the questions.
+We tried to make them as simple as possible using accessibile language, and tried to keep it short as well (@Tourangeau_Rips_Rasinski_2000).
+We also prioritised the size of the survey.
+We kept it to a short 11 questions that will tell us their political standing in the past and present.
+We know what current issues are important to them as well as their age group and the state they are from.
+This will help us gather data without inconveniencing the person.
+
+Finally we take a look at how we sample from the gathered data.
+We thought about random sampling but we were worried about having unequal proportions of the demographic we would advertise to.
+So if more people that go to facebook responded, that doesn't necessairly mean that people use facebook more than the other platforms.
+So we we decided to combine startified and cluster  sampling(@annurev).
+The idea is that we will choose based on the different platforms we advertise to first to have a cluster sampling method.
+And then we will sample the data by dividing the people by the certain aspects of the surevy, like if they are in different age groups or if they are registered voters.
+This is called Stratified Sampling (@annurev).
+Then we take a union of all the samples and we gather a group that we are able to build models based off of.
+
 \newpage

### 2024-11-04 paper/paper.qmd
MODEL VISUALZATION

--- 2024-11-03 paper/paper.qmd
@@
   mutate(
-    end_date_num = as.numeric(end_date - min(end_date)))
-
-harrisdata <- harrisdata |>
-  mutate(state = factor(state))
+    end_date_num = as.numeric(end_date - min(end_date)),
+        state = factor(state))
 
@@
   prior_intercept = priors,
-  weights = harrisdata$numeric_grade,
   seed = 123,
@@
 
-This is Model 2
-
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-spline_model <- stan_glm(
-  pct ~ ns(end_date_num, df = 5) + pollster + state, 
-  data = harrisdata,
-  family = gaussian(),
-  prior = normal(0, 5),
-  prior_intercept = normal(50, 10),
-  seed = 1234,
-  iter = 2000,
-  chains = 4,
-  refresh = 0
-)
-
-# Summarize the model
-summary(spline_model)
-
-# Posterior predictive checks
-pp_check(spline_model)
-
-modelsummary::modelsummary(
-  list(
-    "First model" = spline_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
 ```{r}
@@
 
-harrisdata$pollster <- factor(harrisdata$pollster)
-harrisdata$state <- factor(harrisdata$state)
+harrisdata <- harrisdata %>% drop_na()
+expected_probs <- posterior_epred(bayesian_model, newdata = harrisdata)
+median_probs <- apply(expected_probs, 2, median)
+harrisdata <- harrisdata %>%
+  mutate(median_prob = median_probs)
 
-new_data <- data.frame(
-  end_date_num = seq(
-    min(harrisdata$end_date_num),
-    max(harrisdata$end_date_num),
-    length.out = 100
-  ),
-  pollster = factor("YouGov", levels = levels(harrisdata$pollster)),
-  state = factor("Georgia", levels = levels(harrisdata$state)))
+ggplot(harrisdata, aes(x = end_date_num, y = median_prob, color = state)) +
+  geom_line(alpha = 0.5) +  # Plot raw data with some transparency to show noise
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 10), se = FALSE) +  # Smoother with flexible spline
+  labs(
+    title = "Predicted Probability of Harris Outcome Over Time by State",
+    x = "Days Since First Poll (end_date_num)",
+    y = "Predicted Probability"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~ state, scales = "free_y")
 
-posterior_preds <- posterior_predict(spline_model, newdata = new_data)
+ggplot(harrisdata, aes(x = end_date_num, y = median_prob, color = state)) +
+  geom_line(alpha = 0.5) +  # Plot raw data with some transparency
+  geom_smooth(method = "loess", span = 0.3, se = FALSE) +  # Add a less smoothed trend line
+  labs(
+    title = "Predicted Probability of Harris Outcome Over Time by State",
+    x = "Days Since First Poll (end_date_num)",
+    y = "Predicted Probability"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~ state, scales = "free_y")
 
-pred_summary <- new_data |>
-  mutate(
-    pred_mean = colMeans(posterior_preds),
-    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
-    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
-  )
-
-# Plot data and Bayesian model predictions with confidence ribbon
-ggplot(harrisdata, aes(x = end_date_num, y = pct, color = state)) +
-  geom_point() +
-  geom_line(
-    data = pred_summary,
-    aes(x = end_date_num, y = pred_mean),
-    color = "blue",
-    inherit.aes = FALSE
-  ) +
-  geom_ribbon(
-    data = pred_summary,
-    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
-    alpha = 0.2,
-    inherit.aes = FALSE
-  ) +
-  labs(
-    x = "Days since earliest poll",
-    y = "Percentage",
-    title = "Poll Percentage over Time"
-  ) +
-  facet_wrap(~ state, scales = "free_y") +
-  theme_minimal()
-```
-
-# TRUMP #
-
-```{r}
-trumpdata <- preddata %>%
-  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
-  filter(candidate_name == "Donald Trump") %>%
-  mutate(end_date = mdy(end_date)) %>%
-  filter(end_date >= as.Date("2024-07-21")) %>%
-  mutate(num_trump = round((pct / 100) * sample_size, 0))
-
-ggplot(trumpdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
-  geom_boxplot() +
-  coord_flip() +  # Flip for better readability
-  labs(x = "State", y = "Support Percentage") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-trumpdata <- trumpdata %>%
-  mutate(
-    end_date_num = as.numeric(end_date - min(end_date)))
-
-trumpdata <- trumpdata |>
-  mutate(state = factor(state))
-
-# Define Model
-model2 <- cbind(num_trump, sample_size - num_trump) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
-
-priors <- normal(0, 2.5, autoscale = TRUE)
-
-bayesian_model2 <- stan_glmer(
-  formula = model2,
-  data = trumpdata,
-  family = binomial(link = "logit"),
-  prior = priors,
-  prior_intercept = priors,
-  weights = trumpdata$numeric_grade,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-# Plot checks for Model
-pp_check(bayesian_model2)
-plot(bayesian_model2, pars = "(Intercept)", prob = 0.95)
-modelsummary::modelsummary(
-  list(
-    "First model" = bayesian_model2
-  ),
-  statistic = "mad",
-  fmt = 2
-)
 ```
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-```{r}
-harrisdata <- harrisdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_harris))
-trumpdata <- trumpdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_trump))
-trumpdata
-```
-
-```{r}
-library(dplyr)
-library(ggplot2)
-library(rstanarm)
-
-harris_data <- harrisdata %>%
-  mutate(
-    candidate_name = "Kamala Harris",
-    support_ratio = num_harris  # Assuming harris_support_ratio is already calculated
-  ) %>%
-  select(end_date_num, candidate_name, support_ratio)
-
-# Prepare Trump data with a unified structure
-trump_data <- trumpdata %>%
-  mutate(
-    candidate_name = "Donald Trump",
-    support_ratio = num_trump  # Assuming trump_support_ratio is already calculated
-  ) %>%
-  select(end_date_num, candidate_name, support_ratio)
-
-# Combine both into a single data frame
-combined_data <- bind_rows(harris_data, trump_data)
-
-library(ggplot2)
-library(ggplot2)
-
-# Assuming `combined_data` contains columns `end_date_num`, `support_ratio`, and `candidate_name`
-
-ggplot(combined_data, aes(x = end_date_num, y = support_ratio, color = candidate_name, fill = candidate_name)) +
-  # Shaded area around each line for overlap visualization
-  geom_ribbon(data = combined_data %>% filter(candidate_name == "Kamala Harris"),
-              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
-              alpha = 0.3, fill = "red") +
-  geom_ribbon(data = combined_data %>% filter(candidate_name == "Donald Trump"),
-              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
-              alpha = 0.3, fill = "blue") +
-
-  # Line plot for each candidate
-  geom_line(aes(group = candidate_name), size = 1) +
-
-  scale_color_manual(values = c("Kamala Harris" = "red", "Donald Trump" = "blue")) +  # Set line colors
-  labs(
-    x = "Days Since Earliest Poll", 
-    y = "Support Ratio",
-    color = "Candidate",
-    title = "Support Ratio for Harris and Trump Over Time with Overlap Highlighted"
-  ) +
-  theme_minimal() +
-  theme(
-    text = element_text(size = 12),
-    plot.title = element_text(hjust = 0.5),
-    legend.position = "top",
-    legend.title = element_text(face = "bold")
-  )
-
-```
 # Discussion

### 2024-11-04 paper/paper.qmd
model description

@@
 
-This is Model 1, follows Bayesian Model
-
 We define our model as:
 
-*define each vairable here*
 ```{=tex}
@@
 y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \beta_0 + \beta_1 \times \mbox{State}_i + \beta_2 \times \mbox{Days Since First Poll}_i + \beta_3 \times \mbox{Pollster}_i \\
+\mu_i &= \beta_0 + \beta_1 \times \mbox{State}_i + \beta_2 \times \mbox{Pollster}_i + \beta_3 \times \mbox{Days since first poll}_i\\
 \beta_0 &\sim \mbox{Normal}(0, 2.5) \\
@@
 ```
+
+We used the `stan_glmer` function from `rstanarm` package to create a Bayesian regression model with a Normal distribution.
+The dependent variable in the proportion of respondent who support Kamala Harris, and our model aims to predict Harris' support based on several importnat factors, modeled by:
 
@@
   prior_intercept = priors,
+  # weights = harrisdata$numeric_grade,
   seed = 123,
@@
 ```
+
+where:
+
+- $y_i$ is the dependent variable, represneting the proportion of respondents who support Harris
+- $\beta_0$ is the intercept term, representing the expected proportion of $y_i$ when all other predictors are set to zero.
+- $\beta_1$ corresponds to the `state` choosing one of seven states,
+- $\beta_2$ corresponds to the `pollster`,
+- $\beta_3$ is the value representing number of days since the first poll.
+
+All variables follow the normal distribution with a mean of 0 and a standard deviation of 2.5.
 

### 2024-11-04 paper/paper.qmd
scirpts done and citation done

--- 2024-11-03 paper/paper.qmd
@@
 Each poll has two entries, one giving the Data for the polling opinions of Donald Trump and the second for Kamala Harris.
-We have acquired our polling data from @bibnotmadeyet.
+We have acquired our polling data from @fivethirtyeight2024
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+We use the statistical programming language R [@citeR].
+Our data comes from [@fivethirtyeight2024] and was cleaned, modeled and graphed, using @broom, @tidy, @rstanarm, @janitor, @lubr, @broom, @modelsummary, @splines.
 

### 2024-11-04 scripts/00-simulate_data.R

@@ -1,52 +1,42 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
+# Purpose: Simulates a dataset of American election
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+
 
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
+# Load necessary library
+library(dplyr)
 
+# Set seed for reproducibility
+set.seed(123)
 
-#### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+# Parameters for the simulation
+num_polls <- 100
+pollsters <- c("YouGov", "New York Times", "Ipsos", "The Washington Post", "1892 Polling")
+states <- c("Arizona", "Georgia", "Wisconsin", "Nevada", "Pennsylvania", "North Carolina", "Michigan")
+candidates <- c("Donald Trump", "Kamala Harris", "Jill Stein", "Chase Oliver")
+
+# Generate simulated data
+poll_data <- data.frame(
+  pollster = sample(pollsters, num_polls, replace = TRUE),
+  sample_size = sample(c(800, 1000, 1200, 1500, 2000), num_polls, replace = TRUE),
+  state = sample(states, num_polls, replace = TRUE),
+  candidate_name = sample(candidates, num_polls, replace = TRUE),
+  pct = round(runif(num_polls, 40, 60), 1),  # Simulate percentages between 40 and 60
+  start_date = sample(seq(as.Date('2024-10-01'), as.Date('2024-10-10'), by="days"), num_polls, replace = TRUE),
+  end_date = sample(seq(as.Date('2024-10-11'), as.Date('2024-10-20'), by="days"), num_polls, replace = TRUE),
[…]

### 2024-11-04 scripts/01-test_simulated_data.R

@@ -1,89 +1,55 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
+# Purpose: Tests the structure and validity of the simulated American election data
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
 
+
+
+install.packages("validate")
+install.packages("testthat")
 
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+# Load necessary libraries
+# Load necessary libraries
+library(testthat)
+library(validate)
+library(readr)
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
+# Load the simulated data
+poll_data <- read_csv("data/00-simulated_data/simulated_data.csv")
 
+# Define validation rules
+rules <- validator(
+  n_polls = nrow(poll_data) == 100,
+  n_cols = ncol(poll_data) == 9,
+  is_character_pollster = all(sapply(poll_data$pollster, is.character)),
+  is_numeric_sample_size = all(sapply(poll_data$sample_size, is.numeric)),
+  is_character_state = all(sapply(poll_data$state, is.character)),
+  is_character_candidate = all(sapply(poll_data$candidate_name, is.character)),
+  is_numeric_pct = all(sapply(poll_data$pct, is.numeric)),
+  is_date_start = all(sapply(poll_data$start_date, is.Date)),
+  is_date_end = all(sapply(poll_data$end_date, is.Date)),
+  is_numeric_pollscore = all(sapply(poll_data$pollscore, is.numeric)),
+  is_numeric_grade = all(sapply(poll_data$numeric_grade, is.numeric)),
[…]

### 2024-11-04 scripts/02-download_data.R

@@ -1,26 +1,26 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from [data](https://projects.fivethirtyeight.com/polls/president-general/2024/national/)
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
+# Pre-requisites: go to the site above and search for “Download the data”, then select Presidential general election polls (current cycle), then “Download”
 
 
-#### Workspace setup ####
-library(opendatatoronto)
+
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# Load necessary libraries
+library(testthat)
+library(validate)
+library(readr)
 
 
+# Load the simulated data
+#the data is already downloaded in the repository. 
+data <- read.csv("../data/01-raw_data/raw_data.csv")
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+
+
 
          

### 2024-11-04 scripts/03-clean_data.R

@@ -1,44 +1,27 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Cleans the raw election data by filtering and deleting rows
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
+library(janitor)
+
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+data <- read.csv("./data/01-raw_data/raw_data.csv") |>
+  clean_names()
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+preddata <- data %>%
+  filter(numeric_grade >1.5 & transparency_score >6) %>%
[…]

### 2024-11-04 scripts/04-test_analysis_data.R

@@ -1,69 +1,47 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Tests the data set for any out of place values
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 
 
 #### Workspace setup ####
 library(tidyverse)
+
+# Load necessary libraries
+# Load necessary libraries
 library(testthat)
+library(validate)
+library(readr)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+# Load the simulated data
+poll_data <- read.csv("./data/01-raw_data/raw_data.csv")
 
+# Define validation rules
+rules <- validator(
+  n_polls = nrow(poll_data) == 100,
+  n_cols = ncol(poll_data) == 9,
+  is_character_pollster = all(sapply(poll_data$pollster, is.character)),
+  is_numeric_sample_size = all(sapply(poll_data$sample_size, is.numeric)),
+  is_character_state = all(sapply(poll_data$state, is.character)),
+  is_character_candidate = all(sapply(poll_data$candidate_name, is.character)),
+  is_numeric_pct = all(sapply(poll_data$pct, is.numeric)),
+  is_date_start = all(sapply(poll_data$start_date, is.Date)),
+  is_date_end = all(sapply(poll_data$end_date, is.Date)),
+  is_numeric_pollscore = all(sapply(poll_data$pollscore, is.numeric)),
+  is_numeric_grade = all(sapply(poll_data$numeric_grade, is.numeric)),
+  pollster_in_set = all(poll_data$pollster %in% c("YouGov", "New York Times", "Ipsos", "The Washington Post", "1892 Polling")),
+  sample_size_in_set = all(poll_data$sample_size %in% c(800, 1000, 1200, 1500, 2000)),
[…]

### 2024-11-04 scripts/05-exploratory_data_analysis.R

@@ -13,25 +13,45 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+preddata< - read_csv(preddata, "./data/02-analysis_data/cleaned.csv")
+
+harrisdata <- preddata %>%
+  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  mutate(end_date = mdy(end_date)) %>%
+  filter(end_date >= as.Date("2024-07-21")) %>%
+  mutate(num_harris = round((pct / 100) * sample_size, 0))
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+harrisdata <- harrisdata %>%
+  mutate(
+    end_date_num = as.numeric(end_date - min(end_date)))
+
+harrisdata <- harrisdata |>
+  mutate(state = factor(state))
+
+# Define Model
+model <- cbind(num_harris, sample_size - num_harris) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
+
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model <- stan_glmer(
+  formula = model,
+  data = harrisdata,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  weights = harrisdata$numeric_grade,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95)
+
 
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  bayesian_model,
+  file = "models/bay_model.rds"
 )
 
 

### 2024-11-04 scripts/06-model_data.R

@@ -13,25 +13,50 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+preddata< - read_csv(preddata, "./data/02-analysis_data/cleaned.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+trumpdata <- preddata %>%
+  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
+  filter(candidate_name == "Donald Trump") %>%
+  mutate(end_date = mdy(end_date)) %>%
+  filter(end_date >= as.Date("2024-07-21")) %>%
+  mutate(num_trump = round((pct / 100) * sample_size, 0))
+
+ggplot(trumpdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
+  geom_boxplot() +
+  coord_flip() +  # Flip for better readability
+  labs(x = "State", y = "Support Percentage") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+trumpdata <- trumpdata %>%
+  mutate(
+    end_date_num = as.numeric(end_date - min(end_date)))
+
+trumpdata <- trumpdata |>
+  mutate(state = factor(state))
+
+# Define Model
+model2 <- cbind(num_trump, sample_size - num_trump) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
+
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+bayesian_model2 <- stan_glmer(
+  formula = model2,
+  data = trumpdata,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  weights = trumpdata$numeric_grade,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95)
 
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  bayesian_model2,
+  file = "models/bayesian_model2.rds"
 )
 
 

### 2024-11-04 paper/paper.qmd
merging affected appendix 2

@@
 
-=======
->>>>>>> eda40e1517683f0ad12e0bcd901797a2bb2abd7d
-\newpage
+Here we will be talking about how we would conduct a $100,000 surevey to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent.
+Out of the \$100,000 we will allocate \$15000 for the necessary development and administration of the data.
+This means we will be spending this money to create the survey, and have a infrastructure in place to hold the large amount of data as well as cover any security fees to keep this data private and safe.
+Then we allocate $50000 to advertise the survey.
+We will need a large sample so we will be spending most of our budget for this.
+We will have a different urls for each advertisement we have so that we can incorporate this in our data to see which demographic is accessing the survey through which platform.
+The platforms to advertise this is through spotify, facebook, instagram, and some news networks.
+If possible we will try to use this money to get it endorsed by branches of government to show our reliability.
+Reliability is also what citizens respond to so this will get a lot of respondants.
+We will spend \$25,000 for modeling our data as they tend to be expensive with data this size.
+And the rest of the \$15,000 will be leftover cost for anything that we don't foresee.
+If there is any leftover we can add a survey participation price such as having a free 3 month trial for the platforms we mentioned above (that is if they have a subscription based membership).
+
+We will not be using telephone surveys because according to reserach [@idSurvey] we find the most people, don't pick up calls from unknown numbers and even the people that pick it up they are less inclined to answer the questions of the survey.
+
+Next to look at the actual contents of the survey.
+They can be accessed through this link:[Sample survey question](https://forms.office.com/Pages/ResponsePage.aspx?id=JsKqeAMvTUuQN7RtVsVSEL7G89G56QlHp_qqIPzGbJdUME9UQzJaWUlCVldZTktRV09UQ0Q3N0hQRi4u).
+There are 3 things that we were careful of when we were creating this survey.
+First thing we considered was transperency.
+People need the reassurence that the data that is being collected will not be used against them, and so we teel them what data we do collect and how the data we collect cannot be used to identify a person.
+The second thing we focus on is readability of the questions.
+We tried to make them as simple as possible using accessibile language, and tried to keep it short as well (@Tourangeau_Rips_Rasinski_2000).
+We also prioritised the size of the survey.
+We kept it to a short 11 questions that will tell us their political standing in the past and present.
+We know what current issues are important to them as well as their age group and the state they are from.
+This will help us gather data without inconveniencing the person.
+
+Finally we take a look at how we sample from the gathered data.
+We thought about random sampling but we were worried about having unequal proportions of the demographic we would advertise to.
+So if more people that go to facebook responded, that doesn't necessairly mean that people use facebook more than the other platforms.
+So we we decided to combine startified and cluster  sampling(@annurev).
+The idea is that we will choose based on the different platforms we advertise to first to have a cluster sampling method.
+And then we will sample the data by dividing the people by the certain aspects of the surevy, like if they are in different age groups or if they are registered voters.
+This is called Stratified Sampling (@annurev).
+Then we take a union of all the samples and we gather a group that we are able to build models based off of.
 

### 2024-11-04 paper/paper.qmd
added names and title

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Will Kamala Harris will win the 2024 Presidential election"
+subtitle: "Analysing the polls using baysian modeling"
 author: 
-  - First author
-  - Another author
+  - Yan Mezhiborsky
+  - Cristina Burca
+  - Shreya Sakura Noskor
 thanks: "Code and data are available at: [https://github.com/Mezhi18/US_Election2024.git](https://github.com/Mezhi18/US_Election2024.git)."
@@
 
-harrisdata <- harrisdata %>% drop_na()
+harrisdata <- harrisdata |> drop_na()
 expected_probs <- posterior_epred(bayesian_model, newdata = harrisdata)

### 2024-11-04 scripts/05-exploratory_data_analysis.R

@@ -1,11 +1,8 @@
 #### Preamble ####
 # Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 
 
 #### Workspace setup ####

### 2024-11-04 scripts/06-model_data.R

@@ -1,12 +1,8 @@
 #### Preamble ####
 # Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Author: Sakura Noskor, Yan Mezhiborsky, Cristina Burca
+# Date: 3 November 2024
+# Contact: cristina.burca@mail.utoronto.ca, sakura.noskor@mail.utoronto.ca,  yan.mezhiborsky@mail.utoronto.ca
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-11-04 paper/paper.qmd
finalized model and predicitons

--- 2024-11-04 paper/paper.qmd
@@
   mutate(num_harris = round((pct / 100) * sample_size, 0))
+
+trumpdata <- preddata %>%
+  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
+  filter(candidate_name == "Donald Trump") %>%
+  mutate(end_date = mdy(end_date)) %>%
+  filter(end_date >= as.Date("2024-07-21")) %>%
+  mutate(num_trump = round((pct / 100) * sample_size, 0))
+
+combined_data <- harrisdata %>%
+  inner_join(trumpdata, by = c("state", "pollster", "sample_size", "end_date")) %>%
+  mutate(
+    harris_percentage = (num_harris / sample_size) * 100,
+    trump_percentage = (num_trump / sample_size) * 100
+  )
+
+time_series_data <- combined_data %>%
+  select(state, harris_percentage, trump_percentage) %>%
+  pivot_longer(cols = c(harris_percentage, trump_percentage),
+               names_to = "candidate",
+               values_to = "support_percentage") %>%
+  mutate(candidate = ifelse(candidate == "harris_percentage", "Harris", "Trump"))
+
+# Plotting the support percentages side-by-side for each state
+plot_data <- combined_data %>%
+  pivot_longer(cols = c(harris_percentage, trump_percentage),
+               names_to = "candidate",
+               values_to = "support_percentage") %>%
+  mutate(candidate = ifelse(candidate == "harris_percentage", "Harris", "Trump"))
+
+# Plotting the support percentages side-by-side for each state
+ggplot(plot_data, aes(x = state, y = support_percentage, fill = candidate)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
+  labs(
+    title = "Support Percentages for Harris and Trump by State",
+    x = "State",
+    y = "Support Percentage"
+  ) +
+  scale_fill_manual(values = c("Harris" = "blue", "Trump" = "red"), name = "Candidate") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
 
@@
 ```
+```{r}
+expected_probs <- posterior_epred(bayesian_model, newdata = harrisdata)
+
+# Calculate the mean predicted probability across all observations (overall predicted support percentage)
+overall_predicted_support <- mean(rowMeans(expected_probs)) * 100
+overall_predicted_support
+
+```
+
+```{r}
+expected_probs <- posterior_epred(bayesian_model, newdata = harrisdata)
+
+# Calculate the median predicted probability for each observation
+median_probs <- apply(expected_probs, 2, median)
+
+# Add median predicted probabilities to `harrisdata`
+harrisdata <- harrisdata %>%
+  mutate(median_prob = median_probs)
+
+# Calculate the average predicted support by state
+predicted_support_table <- harrisdata %>%
+  group_by(state) %>%
+  summarize(average_predicted_support = mean(median_prob) * 100)
+
+# Print the table
+predicted_support_table
+```
 

### 2024-11-04 paper/paper.qmd
merge affected appendix 2 AGAIn

--- 2024-11-04 paper/paper.qmd
@@
 # Appendix 2
-
 Here we will be talking about how we would conduct a $100,000 surevey to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent.

### 2024-11-04 paper/paper.qmd
Discussion update

@@
 #| message: false
+#| label: fig-pollingharris
+#| fig-cap: "Polling for Harris By State"
 
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Arizona is a very important swing state especially as the state has switches which party they voted for in in the last two elections with Trump winning 49% of the vote and Clinton with 45.5%, which considering Arizona is a swing state is a rather significant6 margin.
+While in the 2020 election Biden ended up winning Arizona with 49.4% of the vote and Trump 49.1%, which is a much smaller margin of victory and what makes Arizona so questionable in the 2024 election.
+We can see in @fig-pollingharris that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls.
+Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled.
+
+We believe Arizona will vote: **IDK**
 
@@
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%.
+While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992.
+Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, which could very realistically result in a Trump victory in the state in the upcomming election.
 

### 2024-11-04 paper/paper.qmd
.

@@
 While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992.
-Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, which could very realistically result in a Trump victory in the state in the upcomming election.
+Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.
+
+In @fig-pollingharris at the beginning of the polling data the support for Harris was relatively low, while it did start to grow the polling percentage somewhat plateaued at around 47.5% and does not show any signs of growing.
+Furthermore, in @fig-supportstate we can see that Trump has the edge over Harris considering all of these factors there is a very real possibility that Trump does win Georgia at the end of the day.
+
+We believe Georgia will vote: **Trump**
 
 ## Michigan {#sec-michigan}
+
+Harris is currently leading the polls in Michigan with 48% of the vote and Trump with 43%, this is a rather significant margin.
 

### 2024-11-04 paper/paper.qmd
More discussion

@@
 
-As our paper is about the 2024 United States federal election and more specifically we are looking at the polls and the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election.
+As our paper is about the 2024 United States federal election and more specifically we are looking at the polling Data and the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election.
 Our original Data set had over 16,000 unique entries from different pollsters, the business or Organization that conducts the poll.
@@
 
-We believe Arizona will vote: **IDK**
+We believe Arizona will vote: **Donald Trump**
 
@@
 
-We believe Georgia will vote: **Trump**
+We believe Georgia will vote: **Donald Trump**
 
@@
 
-Harris is currently leading the polls in Michigan with 48% of the vote and Trump with 43%, this is a rather significant margin.
+Harris is currently leading the polls in Michigan with 49% of the vote and Trump with 47%, this is a rather slim margin, but the first in the discussed states where Harris is leading in the polls.
+In @fig-pollingharris we can see that there was a significant jump for Harris and a small dip in October followed by a gradual increase leading up to the elections.
+When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner.
+Although unlike the previously mentioned states 2016 was the first time the state voted for the republican candidate since 1992.
+Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state.
+
+We believe Michigan will vote for: **Kamala Harris**
 
@@
 
-## Wisconsin {#sec-Wisconsin}
+## Wisconsin {#sec-wisconsin}
 

### 2024-11-04 paper/paper.qmd
overview finished

--- 2024-11-04 paper/paper.qmd
@@
 ggplot(plot_data, aes(x = state, y = support_percentage, fill = candidate)) +
-  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
   labs(
@@
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
+In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin [CITE NYT PAPER].
+As shown in @fig-supportstate, which displays the support percentages for both Harris and Donald Trump side-by-side by state, the data reveals a competitive race, with relatively close support levels for both candidates in most states.
+While Trump has a slight edge in Arizona, Georgia, and North Carolina, Harris maintains a modest lead in Michigan, Nevada, and Pennsylvania.
+
+Moving to @fig-supportharris, which isolates Harris's support percentages across states, we gain insight into the consistency and variability of her support.
+The box plot highlights states like Wisconsin and Pennsylvania, where the smaller interquartile ranges indicate less variability, suggesting that Harris's support is more stable in these regions.
+Conversely, states such as Nevada and Georgia show greater variability, as reflected in the larger spread of the boxes and whiskers, indicating that Harris's support fluctuates more in these areas.
+
+```{r, fig.height= 3, fig.width= 6}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-supportharris
+#| fig-cap:  "Support Percentage for Harris by State"
 
@@
 ```
-We focus on the swing states of the U.S, according to  [cite NYT PAPER]
 
-These are the statistics of each swing state by percetnage by state.
+To explore temporal trends, we introduce the variable `end_date_num`, representing the number of days since Harris declared her candidacy.
+The next figures examine support trends over time in individual states.
 
-```{r}
+```{r, fig.width= 6, fig.height= 3}
 #| echo: false
@@
 #| message: false
-#| label: fig-pollingharris
-#| fig-cap: "Polling for Harris By State"
+#| label: fig-supportWP
+#| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate for Wisconsin and Pennsylvania"
 
-#### Plot data ####
-base_plot <- ggplot(harrisdata, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date") +
-    theme(legend.position = "bottom")
+# Filter for a subset of states to plot separately
+subset1 <- harrisdata %>% filter(state %in% c("Wisconsin", "Pennsylvania"))
 
-base_plot +
-  geom_point(aes(color = state)) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
-
-ggplot(harrisdata, aes(x = end_date, y = pct, color = state)) +
+# Plot the first subset
+ggplot(subset1, aes(x = end_date, y = pct, color = state)) +
   geom_point() +
@@
   facet_wrap(~ state, scales = "free_y") +
-  labs(y = "Harris Percent", x = "Date", title = "Polling Trends for Kamala Harris by State") +
-  theme_minimal()
+  labs(y = "Harris Percent", x = "Date") +
+  theme_minimal() +
+    scale_x_date(
+    date_breaks = "1 month",            # Set major breaks at monthly intervals
+    date_minor_breaks = "1 week",       # Optional: add minor weekly breaks for granularity
+    date_labels = "%b %d"               # Format to show abbreviated month and day (e.g., "Aug 01")
+  ) +
+   theme(legend.position = "none")
 
 ```
+In @fig-supportWP, Pennsylvania and Wisconsin display distinct trends.
+In Pennsylvania, support for Harris has been relatively consistent since early September, stabilizing around 48.5%.
+In contrast, Wisconsin shows a peak in support in mid-September, followed by a gradual decline that stabilizes in October.
+These observations suggest that while Pennsylvania remains a consistent stronghold for Harris, Wisconsin’s voter sentiment may be more susceptible to fluctuations.
+
+```{r, fig.width= 6, fig.height= 3}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-supportAM
+#| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate"
+
+# Filter for the remaining states
+subset2 <- harrisdata %>% filter(state %in% c( "Arizona", "Michigan"))
+
+# Plot the second subset
+ggplot(subset2, aes(x = end_date, y = pct, color = state)) +
+  geom_point() +
+  geom_smooth(se = TRUE) +
+  facet_wrap(~ state, scales = "free_y") +
+  labs(y = "Harris Percent", x = "Date") +
+  theme_minimal() +
+      scale_x_date(
+    date_breaks = "1 month",
+    date_minor_breaks = "1 week",
+    date_labels = "%b %d" 
+  ) +
+   theme(legend.position = "none")
+```
+@fig-supportAM further examines Harris's support trends in Arizona and Michigan.
+Both states exhibit an upward trend in support starting in early October, which may suggest increasing favorability or strategic campaign efforts in these regions.
+This upward trajectory could indicate an opportunity for Harris to solidify her support in these competitive areas.
+
+```{r, fig.width= 7, fig.height= 3}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-supportGNN
+#| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate"
+
+subset3 <- harrisdata %>% filter(state %in% c("Georgia", "North Carolina", "Nevada"))
+
+# Plot the first subset
+ggplot(subset3, aes(x = end_date, y = pct, color = state)) +
+  geom_point() +
+  geom_smooth(se = TRUE) +
+  facet_wrap(~ state, scales = "free_y") +
+  labs(y = "Harris Percent", x = "Date") +
+  theme_minimal() +
+        scale_x_date(
+    date_breaks = "1 month",
+    date_minor_breaks = "1 week",
+    date_labels = "%b %d" 
+  ) +
+   theme(legend.position = "none")
+```
+In @fig-supportGNN, we analyze Georgia, Nevada, and North Carolina.
+Nevada demonstrates the highest variability in Harris's support, which could reflect the state’s dynamic political response to recent campaign efforts.
+Georgia shows a steady upward trend, indicating growing support for Harris, while North Carolina reveals a slight decline since October, suggesting a possible shift in voter support.
+
+Overall, these visualizations illustrate that voter support for Harris varies significantly across states and over time.
+Notably, many states exhibit shifts in mid-September, indicating that voter opinions are sensitive to external factors or campaign dynamics during this period.
+This temporal variable, alongside state-specific characteristics, will inform the predictive model to better understand and forecast support for Harris in key regions.
 
@@
   group_by(state) %>%
-  summarize(average_predicted_support = mean(median_prob) * 100)
+  summarize(average_predicted_support = round(mean(median_prob) * 100,2))
 
@@
 
-Arizona is a very important swing state especially as the state has switches which party they voted for in in the last two elections with Trump winning 49% of the vote and Clinton with 45.5%, which considering Arizona is a swing state is a rather significant6 margin.
-While in the 2020 election Biden ended up winning Arizona with 49.4% of the vote and Trump 49.1%, which is a much smaller margin of victory and what makes Arizona so questionable in the 2024 election.
-We can see in @fig-pollingharris that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls.
-Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled.
-
-We believe Arizona will vote: **IDK**
+If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 
@@
 
-Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%.
-While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992.
-Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, which could very realistically result in a Trump victory in the state in the upcomming election.
+Please don't use these as sub-heading labels - change them to be what your point actually is.
 

### 2024-11-04 paper/paper.qmd

@@
 #| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate for Wisconsin and Pennsylvania"
+#| label: fig-pollingharris
+#| fig-cap: "Polling for Harris By State"
 
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Arizona is a very important swing state especially as the state has switches which party they voted for in in the last two elections with Trump winning 49% of the vote and Clinton with 45.5%, which considering Arizona is a swing state is a rather significant6 margin.
+While in the 2020 election Biden ended up winning Arizona with 49.4% of the vote and Trump 49.1%, which is a much smaller margin of victory and what makes Arizona so questionable in the 2024 election.
+We can see in @fig-pollingharris that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls.
+Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled.
+
+We believe Arizona will vote: **Donald Trump**
 
@@
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%.
+While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992.
+Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.
+
+In @fig-pollingharris at the beginning of the polling data the support for Harris was relatively low, while it did start to grow the polling percentage somewhat plateaued at around 47.5% and does not show any signs of growing.
+Furthermore, in @fig-supportstate we can see that Trump has the edge over Harris considering all of these factors there is a very real possibility that Trump does win Georgia at the end of the day.
+
+We believe Georgia will vote: **Donald Trump**
 
 ## Michigan {#sec-michigan}
+
+Harris is currently leading the polls in Michigan with 49% of the vote and Trump with 47%, this is a rather slim margin, but the first in the discussed states where Harris is leading in the polls.
+In @fig-pollingharris we can see that there was a significant jump for Harris and a small dip in October followed by a gradual increase leading up to the elections.
+When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner.
+Although unlike the previously mentioned states 2016 was the first time the state voted for the republican candidate since 1992.
+Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state.
+
+We believe Michigan will vote for: **Kamala Harris**
 

### 2024-11-04 paper/paper.qmd
overview, measurements done

@@
 
-```{r}
+```{r, fig.height= 3}
 #| echo: false
@@
   geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
-  labs(
-    title = "Support Percentages for Harris and Trump by State",
+  labs(,
     x = "State",
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, size = 8))
 ```
@@
 #| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate for Wisconsin and Pennsylvania"
-#| label: fig-pollingharris
-#| fig-cap: "Polling for Harris By State"
 
@@
 Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
 Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
 Talk way more about it.
@@
 Add graphs, tables and text.
-
 Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

### 2024-11-04 paper/paper.qmd
More Discussion

@@
 While in the 2020 election Biden ended up winning Arizona with 49.4% of the vote and Trump 49.1%, which is a much smaller margin of victory and what makes Arizona so questionable in the 2024 election.
-We can see in @fig-pollingharris that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls.
+We can see in @fig-supportAM that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls.
 Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled.
@@
 
-In @fig-pollingharris at the beginning of the polling data the support for Harris was relatively low, while it did start to grow the polling percentage somewhat plateaued at around 47.5% and does not show any signs of growing.
+In @fig-supportGNN at the beginning of the polling data the support for Harris was relatively low, while it did start to grow the polling percentage somewhat plateaued at around 47.5% and does not show any signs of growing.
 Furthermore, in @fig-supportstate we can see that Trump has the edge over Harris considering all of these factors there is a very real possibility that Trump does win Georgia at the end of the day.
@@
 
+## Nevada {#sec-nevada}
+
+Nevada is a state who's majority is dessert and wilderness, with two large cities being Reno and Las Vegas, as the case is with many states a Majority of democratic votes come from large cities and the majority of republican votes coming from rural areas.
+Nevada has been a democratic state for the last 4 elections with at least a 2% difference in favor of the democrats in the last two elections.
+
+Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the democratic candidates.
+Nevertheless, Harris is seen to have had a gradual decline in the state stabilizing at around 46%, which which seems low in comparison to other states that are looking more favorably at Harris than Trump.
+Still, when we look at @supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the democrats.
+
+We believe Nevada will vote: **Kamala Harris**
+
+## North Carolina {#sec-norcar}
+
 ## Michigan {#sec-michigan}
@@
 Harris is currently leading the polls in Michigan with 49% of the vote and Trump with 47%, this is a rather slim margin, but the first in the discussed states where Harris is leading in the polls.
-In @fig-pollingharris we can see that there was a significant jump for Harris and a small dip in October followed by a gradual increase leading up to the elections.
+In @fig-supportAM we can see that there was a significant jump for Harris and a small dip in October followed by a gradual increase leading up to the elections.
 When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner.
@@
 
-## Nevada {#sec-nevada}
+## Pennsylvania {#sec-penn}
 
-## North Carolina {#sec-norcar}
+Pennsylvania is an interesting for several reasons, the first is that before Tim Walz was announced as Kamala Harris's running mate, one of the speculated candidates was Josh Shapiro, the current governor of Pennsylvania, which presidential candidates often consider from where their potential running mates are from as they tend to help with swing states nevertheless, she did not choose Shapiro as her running mate and this is no longer a factor.
 
-## Michigan {#sec-michigan}
+Second, Similarly to Nevada and Michigan Pennsylvania Voted for Trump in 2016 with a 48.2% to a 47.5% victory, and for Biden in 2020 with a slightly more significant but not overwhelming victory at 50% to 48.8%.
+2016 was the first time Pennsylvania voted for a republican candidate since 1988.
 
-## Pennsylvania {#sec-penn}
+Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
+
+We believe Pennsylvania will vote: **Kamala Harris**
 

### 2024-11-04 paper/paper.qmd
update visualzation section

--- 2024-11-04 paper/paper.qmd
@@
 - `pct`: This tells us the percentage of participants that intend to vote for the selected candidate.
-- `start_date`: The date the pollster began conducting the poll.
 - `end_date`: The date the pollster finished conducting the poll.
@@
 - `num_harris`: The number of participants that intend to vote for Kamala Harris.
-- `end_date_num`: The number of days since the first poll since Harris announced her candidacy.
+- `end_date_num`: The number of days since the first poll since Harris announced her candidacy, calculated using `end_date`.
 
@@
 
+harrisdata <- harrisdata %>%
+  mutate(
+    end_date_num = as.numeric(end_date - min(end_date)),
+        state = factor(state))
+
 trumpdata <- preddata %>%
@@
 
-```{r, fig.height= 3, fig.width= 6}
+```{r, fig.height= 4, fig.width= 6}
 #| echo: false
@@
 
-ggplot(harrisdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
-  geom_boxplot() +
-  coord_flip() +  # Flip for better readability
-  labs(x = "State", y = "Support Percentage") +
+harrisdata <- harrisdata %>%
+  filter(candidate_name == "Kamala Harris") %>%
+  mutate(end_date = as.Date(end_date, format = "%Y-%m-%d"))
+
+# Plot showing the variation in Harris's support percentages by pollster within each state
+ggplot(harrisdata, aes(x = state, y = pct)) +
+  geom_boxplot(aes(fill = state), alpha = 0.6, outlier.shape = NA) +  # Boxplot to show spread by pollster within each state
+  geom_jitter(aes(color = state), width = 0.2, alpha = 0.4) +  # Adds individual points with jitter to show individual poll results
+  labs(
+    x = "State",
+    y = "Support Percentage"
+  ) +
   theme_minimal() +
@@
 
-Overall, these visualizations illustrate that voter support for Harris varies significantly across states and over time.
-Notably, many states exhibit shifts in mid-September, indicating that voter opinions are sensitive to external factors or campaign dynamics during this period.
-This temporal variable, alongside state-specific characteristics, will inform the predictive model to better understand and forecast support for Harris in key regions.
+```{r, fig.width= 6, fig.height= 3}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-supportpollsters
+#| fig-cap: "Variation in Harris's Support Percentage Across Pollsters Over Time"
 
-## Outcome variables
+pollster_summary <- harrisdata %>%
+  group_by(end_date_num) %>%
+  summarize(
+    mean_support = mean(pct, na.rm = TRUE),
+    median_support = median(pct, na.rm = TRUE),
+    lower_quartile = quantile(pct, 0.25, na.rm = TRUE),
+    upper_quartile = quantile(pct, 0.75, na.rm = TRUE)
+  )
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-Talk more about it.
-Talk way more about it.
+ggplot() +
+  # Draw a line for each pollster from first to last poll
+  geom_line(data = harrisdata %>% group_by(pollster) %>% filter(n() > 1), 
+            aes(x = end_date_num, y = pct, group = pollster, color = pollster), 
+            alpha = 0.1) +
+  # Draw points for each poll (including those with only one poll)
+  geom_point(data = harrisdata, aes(x = end_date_num, y = pct, color = pollster), 
+             size = 0.5, alpha = 0.3) +
+  # Add the median line with ribbon for variability
+  geom_line(data = pollster_summary, aes(x = end_date_num, y = median_support), 
+            color = "blue", size = 1) +
+  geom_ribbon(data = pollster_summary, aes(x = end_date_num, ymin = lower_quartile, ymax = upper_quartile), 
+              fill = "lightblue", alpha = 0.2) +
+  labs(
+    x = "Days Since Declaration",
+    y = "Support Percentage"
+  ) +
+  theme_minimal() +
+  theme(legend.position = "none")
 
-## Predictor variables
+```
+In @fig-supportpollsters, we examine the impact of pollster variability on Harris's support percentage over time.
+This visualization introduces the variable pollster, highlighting the differences in polling results that arise from varying methodologies.
+The chart reveals considerable fluctuations and occasional outliers in Harris' support, likely due to the distinct approaches each pollster employs in sampling and weighting responses.
 
-Add graphs, tables and text.
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+Overall, these visualizations illustrate that Harris’s voter support is not only state-dependent but also fluctuates significantly over time.
+In particular, many states exhibit notable shifts around mid-September, pointing to potential influences from external events or campaign dynamics during this period.
+Furthermore, the observed variability across pollsters emphasizes the necessity of a robust model that can account for differences in pollster methodologies.
+By incorporating temporal patterns, state-specific characteristics, and pollster variability, our predictive model will aim to provide a more comprehensive and nuanced forecast of support for Harris in key regions, helping to isolate genuine shifts in public opinion from noise introduced by polling differences.
+
+## Outcome and Predictor variables
+
+In this analysis, our primary outcome variable is the support percentage for Kamala Harris.
+This variable represents the proportion of respondents in each poll who indicated support for Harris.
+Given the competitive nature of the election, understanding how this support varies across states and over time is crucial for identifying trends and forecasting her overall performance.
+
+To introduce the outcome and predictor variables effectively, you’ll want to set the stage for your model by clearly defining what you are trying to predict (the outcome variable) and the factors that you believe influence it (the predictor variables).
+This section should provide context on why each variable is included and how it might impact the results, without yet delving into the specific mechanics of the model itself.
+
+Here’s an example of how to structure this paragraph:
+
+In this analysis, our primary outcome variable is the support percentage for Kamala Harris.
+This variable represents the proportion of respondents in each poll who indicated support for Harris.
+Given the competitive nature of the election, understanding how this support varies across states and over time is crucial for identifying trends and forecasting her overall performance.
+
+To explain the outcome, we consider several predictor variables:
+
+1.
+State: Since support for candidates can vary significantly across different states due to regional demographics, political history, and local issues, the state variable allows us to capture these geographical differences.
+As shown in the earlier visualizations, some states demonstrate consistently higher or lower support for Harris, emphasizing the importance of including state-level distinctions in our analysis.
+
+2.
+Days Since Declaration: Representing the number of days since Harris declared her candidacy, this temporal variable helps us capture shifts in voter sentiment over time.
+Our analysis reveals several critical points where support trends noticeably change, such as the upward or downward shifts in mid-September.
+This predictor allows the model to account for these temporal patterns and track how Harris’s support evolves as the campaign progresses.
+
+3.
+Pollster: Given that different pollsters employ varied methodologies, sampling techniques, and weighting schemes, the pollster variable captures the potential variability introduced by each polling organization.
+As illustrated in @fig-supportpollsters, the results can vary considerably depending on the pollster, with some reporting notably higher or lower support percentages for Harris.
+By including pollster as a predictor, we aim to control for this source of variability, helping the model focus on underlying trends rather than discrepancies introduced by polling differences.
+
+Each of these predictors provides a distinct layer of insight into Harris’s support trends.
+Together, they form the foundation of our predictive model, which will use these variables to more accurately forecast shifts in voter support and clarify the impact of state, time, and pollster methodology on public opinion.
 
@@
 We used the `stan_glmer` function from `rstanarm` package to create a Bayesian regression model with a Normal distribution.
-The dependent variable in the proportion of respondent who support Kamala Harris, and our model aims to predict Harris' support based on several importnat factors, modeled by:
+The dependent variable in the proportion of respondent who support Kamala Harris, and our model aims to predict Harris' support based on several important factors, modeled by:
 
@@
 #| message: false
-
-# Turn end_date into quantifiable column, to use for model
-harrisdata <- harrisdata %>%
-  mutate(
-    end_date_num = as.numeric(end_date - min(end_date)),
-        state = factor(state))
 

### 2024-11-04 paper/paper.qmd
big update for model and results

@@
 
-## Measurement/ Data Visualization
+## Measurement/ Data Visualization {#sec-graphs}
 
@@
 
-```{r, fig.height= 3}
+```{r, fig.height= 2.5}
 #| echo: false
@@
 harrisdata <- harrisdata %>%
-  mutate(
-    end_date_num = as.numeric(end_date - min(end_date)),
+  mutate(end_date_num = as.numeric(end_date - min(end_date)),
         state = factor(state))
@@
     harris_percentage = (num_harris / sample_size) * 100,
-    trump_percentage = (num_trump / sample_size) * 100
-  )
+    trump_percentage = (num_trump / sample_size) * 100)
 
@@
 
-# Plotting the support percentages side-by-side for each state
 plot_data <- combined_data %>%
@@
 
-# Plotting the support percentages side-by-side for each state
 ggplot(plot_data, aes(x = state, y = support_percentage, fill = candidate)) +
@@
   ) +
-  scale_fill_manual(values = c("Harris" = "blue", "Trump" = "red"), name = "Candidate") +
+  scale_fill_manual(values = c("Harris" = "deepskyblue2", "Trump" = "tomato1"), name = "Candidate") +
   theme_minimal() +
@@
 
-```{r, fig.height= 4, fig.width= 6}
+```{r, fig.height= 3.5, fig.width= 6.5}
 #| echo: false
@@
 
-# Plot showing the variation in Harris's support percentages by pollster within each state
 ggplot(harrisdata, aes(x = state, y = pct)) +
-  geom_boxplot(aes(fill = state), alpha = 0.6, outlier.shape = NA) +  # Boxplot to show spread by pollster within each state
-  geom_jitter(aes(color = state), width = 0.2, alpha = 0.4) +  # Adds individual points with jitter to show individual poll results
+  geom_boxplot(aes(fill = state), alpha = 0.6, outlier.shape = NA) + 
+  geom_jitter(aes(color = state), width = 0.2, alpha = 0.4) + 
   labs(
     x = "State",
-    y = "Support Percentage"
-  ) +
+    y = "Support Percentage") +
   theme_minimal() +
@@
 
-```{r, fig.width= 6, fig.height= 3}
+```{r, fig.width= 6, fig.height= 2.5}
 #| echo: false
@@
 
-# Filter for a subset of states to plot separately
 subset1 <- harrisdata %>% filter(state %in% c("Wisconsin", "Pennsylvania"))
 
-# Plot the first subset
 ggplot(subset1, aes(x = end_date, y = pct, color = state)) +
@@
     scale_x_date(
-    date_breaks = "1 month",            # Set major breaks at monthly intervals
-    date_minor_breaks = "1 week",       # Optional: add minor weekly breaks for granularity
-    date_labels = "%b %d"               # Format to show abbreviated month and day (e.g., "Aug 01")
-  ) +
+    date_breaks = "1 month",           
+    date_minor_breaks = "1 week",     
+    date_labels = "%b %d" ) +
    theme(legend.position = "none")
@@
 ```
+
 In @fig-supportWP, Pennsylvania and Wisconsin display distinct trends.
@@
 
-```{r, fig.width= 7, fig.height= 3}
+```{r, fig.width= 7, fig.height= 2.5}
 #| echo: false
@@
 ```
+
 In @fig-supportGNN, we analyze Georgia, Nevada, and North Carolina.
@@
 
-```{r, fig.width= 6, fig.height= 3}
+```{r, fig.width= 5, fig.height= 2.5}
 #| echo: false
@@
   # Draw a line for each pollster from first to last poll
-  geom_line(data = harrisdata %>% group_by(pollster) %>% filter(n() > 1), 
+  geom_point(data = harrisdata %>% group_by(pollster) %>% filter(n() > 1), 
             aes(x = end_date_num, y = pct, group = pollster, color = pollster), 
@@
   geom_line(data = pollster_summary, aes(x = end_date_num, y = median_support), 
-            color = "blue", size = 1) +
+            color = "royalblue", size = 1) +
   geom_ribbon(data = pollster_summary, aes(x = end_date_num, ymin = lower_quartile, ymax = upper_quartile), 
@@
 ```
+
 In @fig-supportpollsters, we examine the impact of pollster variability on Harris's support percentage over time.
@@
 
-To introduce the outcome and predictor variables effectively, you’ll want to set the stage for your model by clearly defining what you are trying to predict (the outcome variable) and the factors that you believe influence it (the predictor variables).
-This section should provide context on why each variable is included and how it might impact the results, without yet delving into the specific mechanics of the model itself.
-
-Here’s an example of how to structure this paragraph:
-
 In this analysis, our primary outcome variable is the support percentage for Kamala Harris.
@@
 1.
-State: Since support for candidates can vary significantly across different states due to regional demographics, political history, and local issues, the state variable allows us to capture these geographical differences.
+**State**: Since support for candidates can vary significantly across different states due to regional demographics, political history, and local issues, the state variable allows us to capture these geographical differences.
 As shown in the earlier visualizations, some states demonstrate consistently higher or lower support for Harris, emphasizing the importance of including state-level distinctions in our analysis.
@@
 2.
-Days Since Declaration: Representing the number of days since Harris declared her candidacy, this temporal variable helps us capture shifts in voter sentiment over time.
+**Days Since Declaration**: Representing the number of days since Harris declared her candidacy, this temporal variable helps us capture shifts in voter sentiment over time.
 Our analysis reveals several critical points where support trends noticeably change, such as the upward or downward shifts in mid-September.
@@
 3.
-Pollster: Given that different pollsters employ varied methodologies, sampling techniques, and weighting schemes, the pollster variable captures the potential variability introduced by each polling organization.
+**Pollster**: Given that different pollsters employ varied methodologies, sampling techniques, and weighting schemes, the pollster variable captures the potential variability introduced by each polling organization.
 As illustrated in @fig-supportpollsters, the results can vary considerably depending on the pollster, with some reporting notably higher or lower support percentages for Harris.
@@
 
-We define our model as:
+## Model Step- Up
+
+We used the `stan_glmer` function from the `rstanarm` package to create a Bayesian regression model with a binomial family and logit link.
+This model is designed to predict the proportion of respondents who support Kamala Harris based on several key factors.
+
+The model specification is as follows:
 
@@
 \beta_2 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_3 &\sim \mbox{Exponential}(1)
+\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
 \end{align}
@@
 
-We used the `stan_glmer` function from `rstanarm` package to create a Bayesian regression model with a Normal distribution.
-The dependent variable in the proportion of respondent who support Kamala Harris, and our model aims to predict Harris' support based on several important factors, modeled by:
+The outcome variable in this model is the proportion of respondents who support Harris in each poll, represented as $y_i$.
+This binomial outcome (successes out of the total sample size) is modeled as a function of several predictor variables that account for differences by state, pollster, and time.
+
+- $y_i$ is the dependent variable, representing the proportion of respondents who support Harris.
+- $\beta_0$: The intercept term, representing the expected proportion of $y_i$ when all other predictors are set to zero.
+- **State** ($\beta_1$): Categorical variable that accounts for differences in support across seven key swing states.
+- **Pollster** ($\beta_2$): Categorical variable capturing variability introduced by the polling organization.
+- **Days since first poll** ($\beta_3$): Numeric variable representing the number of days since Harris declared her candidacy.
+
+Each coefficient $beta$ follows the normal distribution with a mean of 0 and a standard deviation of 2.5, providing some regularization while allowing for a range of possible effects.
+The standard deviation $\sigma$ is given an Exponential(1) prior, which constrains it to positive values and provides a realistic expectation of variability around the mean
+Each predictor captures a unique dimension of voter sentiment:
+
+- **State** accounts for regional differences in support due to factors like demographics and political history.
+- **Pollster** adjusts for methodological variations in polling, which we observed to influence support percentages.
+- **Days Since Declaration** captures temporal trends, allowing the model to incorporate fluctuations in voter sentiment over the campaign period.
+
+By combining these predictors, the model aims to forecast Harris’s support by isolating genuine trends from random variability.
+
+### Model justification
+
+Based on the data overview in @sec-graphs, we observe that Harris’s support varies across key swing states, and we expect these regional differences to influence our model's predictions.
+The state variable is anticipated to capture these geographic variations, with states like Arizona, Michigan, Georgia, and Pennsylvania showing relatively consistent high levels of support for Harris.
+In contrast, Nevada, North Carolina, and Wisconsin exhibit a declining trend in support.
+We expect the intercept to be around 50%, as this reflects the average level of support observed across all states and pollsters.
+
+A model validation is included in `scripts/05-exploratory_data_analysis.R`, and a Posterior Predictive Check and summary statistics for the model are included in @sec-checks.
+
+## Model Results
+
+The model predicts an average support percentage of approximately 47.33% for Kamala Harris across the selected states.
+This percentage reflects the central tendency of predicted support within the dataset, factoring in the influence of state, pollster, and the number of days since the first poll.
+
+The results indicate variations in predicted support across states.
+For instance, Pennsylvania and Wisconsin show the highest average predicted support percentages (48.02% and 48.38%, respectively), while states like Nevada and Arizona display slightly lower support.
+
+In states like Arizona, Georgia, Michigan, and North Carolina, the model predicts a steady increase in support for Harris over time, as shown in @fig-modelGAWM.
+In states such as Nevada, Pennsylvania, and Wisconsin, the model predicts a decline in support over time, as seen in @fig-modelNPW.
 
 ```{r}
-#| include: false
+#| echo: false
 #| warning: false
 #| message: false
+#| label: fig-modelGAWM
+#| fig-cap: "Poll predictions of Arizona, Geogria, Michigan, North Carolina"
 
-# Define Model
-model <- cbind(num_harris, sample_size - num_harris) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)
+bay_model <-
+  readRDS(file = here::here("models/bay_model.rds"))
 
-priors <- normal(0, 2.5, autoscale = TRUE)
+harrisdata_clean <- harrisdata %>% drop_na()
 
-bayesian_model <- stan_glmer(
-  formula = model,
-  data = harrisdata,
-  family = binomial(link = "logit"),
-  prior = priors,
-  prior_intercept = priors,
-  # weights = harrisdata$numeric_grade,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95)
+harrisdata <- harrisdata %>% drop_na()
+expected_probs <- posterior_epred(bay_model, newdata = harrisdata)
+median_probs <- apply(expected_probs, 2, median)
+harrisdata <- harrisdata %>%
+  mutate(median_prob = median_probs)
+
+subset_up <- harrisdata %>% filter(state %in% c("Arizona", "Georgia", "Michigan", "North Carolina"))
+
+ggplot(subset_up, aes(x = end_date_num, y = median_prob, color = state)) +
+  geom_line(alpha = 0.5) +  
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 10), se = FALSE) +
+  labs(
+    x = "Days Since First Poll",
+    y = "Predicted Probability"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~ state) +
+  theme(legend.position = "none")
+
+```
+In @fig-modelNPW, we see states of Nevada, Pennsylvania and Wisconsin have a singular declining trend.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modelNPW
+#| fig-cap: "Poll predictions of Nevada, Pennsylvania, Wisconsin"
+
+subset_down <- harrisdata %>% filter(state %in% c("Nevada", "Pennsylvania", "Wisconsin"))
+
+ggplot(subset_down, aes(x = end_date_num, y = median_prob, color = state)) +
+  geom_line(alpha = 0.5) +
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 10), se = FALSE) + 
+  labs(
+    x = "Days Since First Poll",
+    y = "Predicted Probability"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~ state, ncol = 2) +
+    theme(legend.position = "none")
+
 ```
 
-where:
-
-- $y_i$ is the dependent variable, represneting the proportion of respondents who support Harris
-- $\beta_0$ is the intercept term, representing the expected proportion of $y_i$ when all other predictors are set to zero.
-- $\beta_1$ corresponds to the `state` choosing one of seven states,
-- $\beta_2$ corresponds to the `pollster`,
-- $\beta_3$ is the value representing number of days since the first poll.
-
-All variables follow the normal distribution with a mean of 0 and a standard deviation of 2.5.
+CONFIDENCE INTERVAL TABLE ??
 
@@
 
-# Plot checks for Model
-pp_check(bayesian_model)
-plot(bayesian_model, pars = "(Intercept)", prob = 0.95)
-modelsummary::modelsummary(
-  list(
-    "First model" = bayesian_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-
-harrisdata <- harrisdata %>% drop_na()
-expected_probs <- posterior_epred(bayesian_model, newdata = harrisdata)
-median_probs <- apply(expected_probs, 2, median)
-harrisdata <- harrisdata %>%
-  mutate(median_prob = median_probs)
-
-ggplot(harrisdata, aes(x = end_date_num, y = median_prob, color = state)) +
-  geom_line(alpha = 0.5) +  # Plot raw data with some transparency to show noise
-  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 10), se = FALSE) +  # Smoother with flexible spline
-  labs(
-    title = "Predicted Probability of Harris Outcome Over Time by State",
-    x = "Days Since First Poll (end_date_num)",
-    y = "Predicted Probability"
-  ) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  facet_wrap(~ state, scales = "free_y")
-
-ggplot(harrisdata, aes(x = end_date_num, y = median_prob, color = state)) +
-  geom_line(alpha = 0.5) +  # Plot raw data with some transparency
-  geom_smooth(method = "loess", span = 0.3, se = FALSE) +  # Add a less smoothed trend line
-  labs(
-    title = "Predicted Probability of Harris Outcome Over Time by State",
-    x = "Days Since First Poll (end_date_num)",
-    y = "Predicted Probability"
-  ) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  facet_wrap(~ state, scales = "free_y")
-
-```
-```{r}
-expected_probs <- posterior_epred(bayesian_model, newdata = harrisdata)
-
-# Calculate the mean predicted probability across all observations (overall predicted support percentage)
-overall_predicted_support <- mean(rowMeans(expected_probs)) * 100
-overall_predicted_support
-
-```
-
-```{r}
-expected_probs <- posterior_epred(bayesian_model, newdata = harrisdata)
+expected_probs <- posterior_epred(bay_model, newdata = harrisdata)
 
@@
 ```
+```{r}
+
+```
 
@@
 
-## Model set-up
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-What do we expect...
-We expect a predictions in the high 40% based on our current data.
-
 # Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
 
@@
 
-# Appendix {.unnumbered}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
+# Appendix 1
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+# Appendix 2
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: false
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
-\newpage
-
-# Appendix 1
-
-# Appendix 2
 Here we will be talking about how we would conduct a $100,000 surevey to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent.
@@
 If possible we will try to use this money to get it endorsed by branches of government to show our reliability.
-Reliability is also what citizens respond to so this will get a lot of respondants.
+Reliability is also what citizens respond to so this will get a lot of respondents.
 We will spend \$25,000 for modeling our data as they tend to be expensive with data this size.
@@
 
+# Appendix 3 - checks and summary stats {#sec-checks}
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Plot checks for Model
+pp_check(bay_model)
+plot(bay_model, pars = "(Intercept)", prob = 0.95)
+modelsummary::modelsummary(
+  list(
+    "First model" = bay_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
+
 # References

### 2024-11-04 paper/paper.qmd
methodologies and measurements added + fix table

@@
 library(rstanarm)
-library(splines)
+library(knitr)
 ```
@@
 
-We use the statistical programming language R [@citeR].
-Our data comes from [@fivethirtyeight2024] and was cleaned, modeled and graphed, using @broom, @tidy, @rstanarm, @janitor, @lubr, @broom, @modelsummary, @splines.
+We use the statistical programming language R [@citeR]. Our data comes from [@fivethirtyeight2024] and was cleaned, modeled and graphed, using @broom, @tidy, @rstanarm, @janitor, @lubr, @modelsummary.
 
-## Methodology
+## Methodology and Measurement
 
-This paper uses 'polls-of-polls' method to analyze and predict our outcome, which combines polls from multiple sources (pollsters).
-By combining results from various sources, this approach incorporates diverse perspectives, which helps to minimize individual biases and provides a more balanced view.
-Unlike relying on a single poll, which can be influenced by its own biases, the 'polls-of-polls' method enhances reliability and increases the overall validity of the results.
+This paper employs a "polls-of-polls" approach to analyze and predict our outcome, combining data from multiple pollsters.
+By integrating results from various sources, this method captures a broader range of perspectives, reducing the influence of individual biases and offering a more balanced view.
+Unlike relying on a single poll, which may be shaped by its unique methodological biases, the "polls-of-polls" approach enhances reliability and increases the overall validity of our predictions.
 
-... explain how pollsters survey people, what their different methodologies are, and why these things are important for an accurate prediction...
+Each pollster uses methodologies they believe are most effective for capturing an accurate snapshot of public opinion.
+Pollsters begin by selecting a sample of voters they consider representative of the broader population, factoring in various demographics such as age, region, and socio-economic background.
+However, polling methods differ significantly; some pollsters rely on phone interviews, while others use online surveys, text messaging, or a combination of methods (mixed-mode).
+After data collection, pollsters apply weighting adjustments to account for demographic imbalances, ensuring that underrepresented groups are adequately reflected in the results.
 
-## Measurement/ Data Visualization {#sec-graphs}
+The diversity in methodologies helps reduce systematic bias compared to using a single poll source.
+Different methods may also be more effective in reaching certain demographics—such as older voters through phone surveys and younger voters through online platforms—broadening the inclusivity and representativeness of the sample.
+
+n predictive modeling, accounting for pollster variability and methodology is essential to avoid overfitting to any single source.
+By including pollster as a predictor in the model, we allow it to recognize that poll results may vary not only due to real changes in public opinion but also because of differences in survey methodologies.
+This consideration enables the model to generate more accurate, generalized predictions, offering a reliable forecast that reflects a well-rounded view of public opinion
+
+## Data Visualization {#sec-graphs}
 
@@
 
-Each coefficient $beta$ follows the normal distribution with a mean of 0 and a standard deviation of 2.5, providing some regularization while allowing for a range of possible effects.
+Each coefficient $\beta$ follows the normal distribution with a mean of 0 and a standard deviation of 2.5, providing some regularization while allowing for a range of possible effects.
 The standard deviation $\sigma$ is given an Exponential(1) prior, which constrains it to positive values and provides a realistic expectation of variability around the mean
@@
   group_by(state) %>%
-  summarize(average_predicted_support = round(mean(median_prob) * 100,2))
+  summarize(average_predicted_support = round(mean(median_prob) * 100,2)) %>%
+    arrange(average_predicted_support)
 
-# Print the table
-predicted_support_table
+#predicted_support_table <- predicted_support_table[, c("state", "average_predicted_support")]
+
+kable(predicted_support_table, col.names = c("State", "Avg. Predicted Support for Harris"), digits = 2)
 ```

### 2024-11-04 paper/paper.qmd
update appendix 2 and readme and abstarcat

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "We analyse the probability of Kamala Harris winning in the 7 swing states of USA using baysian modeling; Arizona, Michigan, Pennsylvania, Nevada, Georgia, Wisconsin, North Carolina.
+USA is a large country with its national economy affecting the global economic conditions, which is why predicting the future President of USA will help understand the future economic condiotion of the world.
+Through baysian modelling we found that North Carolina (47.26% support for Harris), Nevada(46.43% support for Harris), Michigan(47.30% support for Harris), Pennsylvania(48.02% support for Harris) will vote Kamala Harris, while the rest of the swing states might vote Donald Trump, meaning that majority of the swing states will vote Kamala Harris.
+Provided this we predict that Kamal Harris will win the election.
+By creating a baysian model based on \"polls of polls \" where we compare the results from different polls, we are able to make this prediction.
+"
 format:
@@
 
-```{r, fig.width= 6, fig.height= 2.5}
+```{r, fig.width= 6, fig.height= 2}
 #| echo: false
@@
 
-```{r, fig.width= 5, fig.height= 2.5}
+```{r, fig.width= 6, fig.height= 2}
 #| echo: false

### 2024-11-04 paper/paper.qmd
update appendix 2 and readme and abstarcat

@@
 
-```{r, fig.width= 6, fig.height= 2}
+```{r, fig.width= 6, fig.height= 2.5}
 #| echo: false
@@
 
-```{r, fig.width= 6, fig.height= 2}
+```{r, fig.width= 5, fig.height= 2.5}
 #| echo: false

### 2024-11-04 paper/paper.qmd
Minor Edit

--- 2024-11-04 paper/paper.qmd
@@
 abstract: "We analyse the probability of Kamala Harris winning in the 7 swing states of USA using baysian modeling; Arizona, Michigan, Pennsylvania, Nevada, Georgia, Wisconsin, North Carolina.
-USA is a large country with its national economy affecting the global economic conditions, which is why predicting the future President of USA will help understand the future economic condiotion of the world.
+USA is a large country with its national economy affecting the global economic conditions, which is why predicting the future President of the USA will help understand the future economic condiotion of the world.
 Through baysian modelling we found that North Carolina (47.26% support for Harris), Nevada(46.43% support for Harris), Michigan(47.30% support for Harris), Pennsylvania(48.02% support for Harris) will vote Kamala Harris, while the rest of the swing states might vote Donald Trump, meaning that majority of the swing states will vote Kamala Harris.
@@
 
-In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin [CITE NYT PAPER].
+In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin \[CITE NYT PAPER\].
 As shown in @fig-supportstate, which displays the support percentages for both Harris and Donald Trump side-by-side by state, the data reveals a competitive race, with relatively close support levels for both candidates in most states.
@@
 ```
+
 @fig-supportAM further examines Harris's support trends in Arizona and Michigan.
@@
 ```
-
 The outcome variable in this model is the proportion of respondents who support Harris in each poll, represented as $y_i$.
@@
 Each coefficient $\beta$ follows the normal distribution with a mean of 0 and a standard deviation of 2.5, providing some regularization while allowing for a range of possible effects.
-The standard deviation $\sigma$ is given an Exponential(1) prior, which constrains it to positive values and provides a realistic expectation of variability around the mean
-Each predictor captures a unique dimension of voter sentiment:
+The standard deviation $\sigma$ is given an Exponential(1) prior, which constrains it to positive values and provides a realistic expectation of variability around the mean Each predictor captures a unique dimension of voter sentiment:
 
@@
 ```
+
 In @fig-modelNPW, we see states of Nevada, Pennsylvania and Wisconsin have a singular declining trend.
@@
 ```
+
 ```{r}
@@
 
-Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
+Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.\@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
 
@@
 
-Here we will be talking about how we would conduct a $100,000 surevey to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent.
+Here we will be talking about how we would conduct a \$100,000 surevey to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent.
 Out of the \$100,000 we will allocate \$15000 for the necessary development and administration of the data.
 This means we will be spending this money to create the survey, and have a infrastructure in place to hold the large amount of data as well as cover any security fees to keep this data private and safe.
-Then we allocate $50000 to advertise the survey.
+Then we allocate \$50000 to advertise the survey.
 We will need a large sample so we will be spending most of our budget for this.

### 2024-11-04 paper/paper.qmd
intro done

--- 2024-11-04 paper/paper.qmd
@@
 USA is a large country with its national economy affecting the global economic conditions, which is why predicting the future President of USA will help understand the future economic condiotion of the world.
-Through baysian modelling we found that North Carolina (47.26% support for Harris), Nevada(46.43% support for Harris), Michigan(47.30% support for Harris), Pennsylvania(48.02% support for Harris) will vote Kamala Harris, while the rest of the swing states might vote Donald Trump, meaning that majority of the swing states will vote Kamala Harris.
+Through baysian modelling we found that North Carolina (47.26% support for Harris), Nevada(46.43% support for Harris), Wisconsin(48.38% support for Harris), Michigan(47.30% support for Harris), Pennsylvania(48.02% support for Harris) will vote Kamala Harris, while the rest of the swing states might vote Donald Trump, meaning that majority of the swing states will vote Kamala Harris.
 Provided this we predict that Kamal Harris will win the election.
@@
 Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the forces shaping the 2024 presidential election.
+We use Bayesian regression model to create a model to predict the percent of population that will support in the swing states based on pollster, the number of days since Kamala Harris joined the race.
+More on the model and the variables, in the Data and Model sections.
 
-Estimand paragraph
+We noticed in most predictions of the presidential candidate they are usually based on one survey or pollster and they try to gather data from the one poll to predict the the future President.
+However, through looking at the results from different polls we are able to get a better, more accurate prediction.
+We also look at over state polls which allows us to focus on the swing states.
+Swing states are states that change sides every election.
+The 2 major parties that fight for power are the Republicans and the Democrats.
+Some states are always voting for one party regardless of the candidate so, the swing states are the once that tell us more about who might win the election.
+The gap we are trying to cover is predicting the future President of USA by looking at different polls, while being as effecient as possible by looking only at the swing states.
 
+To go a little more in-depth into what we actually did, we have to first discuss our estimand.
+Our estimand, the parameter we aim to study, is the predicted support for Kamala Harris in the swing states (measured in %).
+We do this by creating Baysian model with 3 parameters; state, pollster, and says since Kamala Harris was announced as candidate.
+We have the priors set with a normal distribution centered at 0, with a standard deviation of 2.5.
+We expect that because the swing states are states that don't have a party that they dominantly vote for, the expected support will be around 50%.
+We also make other graphs in the Data section, which helps us understand the values themselves.
+
+From the result section we find that the average predicted support for Kamala Harris for the 7 swing states were as follows: Arizona at 46.26%, Nevada at 46.43%, Georgia at 46.59%, North Carolina at 47.26%, Michigan at 47.30%, Pennsylvania at 48.02%, Wisconsin at 48.38%.
+Based on this we predict that Arizona, and Georgia will vote Donald trump and the rest of the states will vote Kamala Harris.
+
+Predicting who will become president is important because it helps people understand and prepare for future changes in the country’s policies and priorities.
+A new president can shape laws, affect the economy, influence international relationships, and address key issues like healthcare, education, and climate change.
+Accurate predictions allow citizens, businesses, and governments to plan and adapt.
+For example, if polls indicate a particular candidate might win, groups can start thinking about how their plans might align with the candidate's proposed policies.
+It also helps keep people informed and engaged in the political process, making them more likely to vote and participate in democracy.
 Results paragraph
 
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The paper is structured with the introduction in the beginning followed by the Data section wich takes a look at the methodology of the paper and how the data was gathered.
+We explain the variables we isolated in order for this analysis to be complete.
+Then it is followed by the Model section where we introduce the model and explain the graphs of related models.
+The result sections will cover the model summary and summary table of the predictions for each of the swing states.
+The last section is the discussion section where we talk about what the results means and we give our final conclusion and weaknesses.
+We also give a final word about who will win the upcoming election.
+We also have additional appendices, where we talk about the methodology of the polls and how they compare.
+The second appendix talks about a survey we designed with reasons and disadvantage.
+The third appendix just shows the models in more depth.
 
@@
 Next to look at the actual contents of the survey.
-They can be accessed through this link:[Sample survey question](https://forms.office.com/Pages/ResponsePage.aspx?id=JsKqeAMvTUuQN7RtVsVSEL7G89G56QlHp_qqIPzGbJdUME9UQzJaWUlCVldZTktRV09UQ0Q3N0hQRi4u).
+They can be accessed through this link: [Sample survey question](https://forms.office.com/Pages/ResponsePage.aspx?id=JsKqeAMvTUuQN7RtVsVSEL7G89G56QlHp_qqIPzGbJdUME9UQzJaWUlCVldZTktRV09UQ0Q3N0hQRi4u).
 There are 3 things that we were careful of when we were creating this survey.
 First thing we considered was transperency.
-People need the reassurence that the data that is being collected will not be used against them, and so we teel them what data we do collect and how the data we collect cannot be used to identify a person.
+People need the reassurance that the data that is being collected will not be used against them, and so we feel them what data we do collect and how the data we collect cannot be used to identify a person.
 The second thing we focus on is readability of the questions.
@@
 This will help us gather data without inconveniencing the person.
+There are disadvanatages to the way we structured the survey.
+For example, we were keeping time and security in mind, so we restricted the survey to 11 questions, however, this limits the amount of answers we get from the reader.
+We may be missing some crucial information that helps us figure out which group of people may vote for whom, such as ethnicity and wages (@annurev).
 
@@
 We thought about random sampling but we were worried about having unequal proportions of the demographic we would advertise to.
-So if more people that go to facebook responded, that doesn't necessairly mean that people use facebook more than the other platforms.
-So we we decided to combine startified and cluster  sampling(@annurev).
+So if more people that go to facebook responded, that doesn't necessarily mean that people use facebook more than the other platforms.
+So we we decided to combine stratified and cluster  sampling(@annurev).
 The idea is that we will choose based on the different platforms we advertise to first to have a cluster sampling method.
@@
 Then we take a union of all the samples and we gather a group that we are able to build models based off of.
+There are obvious disadvantages as well.
+For example, we have fewer questions meaning when we sample based on the question we might have cases where, some groups are over or under represented.
+This can happen for example when most of the people who answered might be from age 18-50 because those are the age groups with access to a phone and are frequently visiting the sites we advertise our survey on.
+Also the different apps may skew the data as Facebook has a older audience as opposed to Instagram and Tiktok.
+Another disadvantage is the fact that we have 2 different sampling methods that we use and we combine them both, meaning that this can be an expensive process.
+Doing so may exceed the allocated budget however, it will be worth it because we minimize the sampling bias we talk about earlier where we may have a group that has more access to the survey.
+The last disadvantage we also foresee is the fact that younger audience may not be as interested in filling out a survey about politics as they feel, it might not directly affect them.
+We try to minimize this by offering compensation for their time when they fill out the survey, i.e.
+free membership services for 3 months on Spotify.
+
+Although there are both advatages and disadvantages, we feel that this survey is well though out and doesn't allow for ambiguity in the answers of the participants.
 

### 2024-11-04 paper/paper.qmd
porse fixes

--- 2024-11-04 paper/paper.qmd
@@
 ```
-```{r}
-
-```
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
 

### 2024-11-04 paper/paper.qmd
merge confluict resolved

--- 2024-11-04 paper/paper.qmd
@@
 Provided this we predict that Kamal Harris will win the election.
-By creating a baysian model based on \"polls of polls \" where we compare the results from different polls, we are able to make this prediction.
-"
+By creating a baysian model based on \"polls of polls \" where we compare the results from different polls, we are able to make this prediction."
 format:

### 2024-11-04 paper/paper.qmd

--- 2024-11-04 paper/paper.qmd
@@
 #| message: false
+#| label: fig-ppcheck
+#| fig-cap: "Posterior predictive check of model"
 
@@
 pp_check(bay_model)
-plot(bay_model, pars = "(Intercept)", prob = 0.95)
-modelsummary::modelsummary(
-  list(
-    "First model" = bay_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+
 ```
 
+In @fig-ppcheck, we check the Posterior predictive check of model of our model. Since the raw data aligns closely with our prdictions, our model is a good predicton of the data.
+
+\newpage
+
 # References

### 2024-11-04 paper/paper.qmd
finished appendix

@@
 
-Why we choose NYT?
-- `numeric_grade` is 3 - `pollscore` is -1.5, a score of reliability called "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.", where negative numbers are better - `transparency_score` is 9, reflects pollsters transparency about their methodology (calculated based on how much information it discloses about its polls and weighted by recency) - `population_full` is 'rv', respondents are registered voters
+The New York Times/Siena College Poll is a respected polling partnership that provides data-driven insights into U.S.
+public opinion, particularly on political, social, and economic issues.
+Known for its rigorous methodology, this poll focuses on capturing the views of registered and likely voters across the country, with particular emphasis on key battleground states during election cycles.
+The poll is frequently cited for its accuracy and transparency, and is widely regarded as reliable by the public, political analysts, and the media.
+
+The New York Times/Siena College Poll is conducts live telephone interviews with a national sample of registered voters, using both landlines and cellphones, with over 90% of contacts made via cellphones.
+The sample frame is drawn from a comprehensive voter file covering 200 million registered voters.
+This file provides demographic and geographic information, enabling representative sampling across age, race, and political affiliation.
+The poll targets registered U.S.
+voters, with special focus on battleground states like Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin, determined as swing states.
+Respondents are randomly selected from the voter file, ensuring a diverse sample by region, party, and demographics.
+To mitigate non-response bias, the poll applies demographic weighting, increasing the weight of underrepresented groups, such as young adults or non-college graduates.
+Low response rates (around 2%) present challenges, to mitigate this, weighing adjustments are applied to correct for demographic imbalances and potential non-response bias.
+Questions are structured to be clear, neutral, and representative across political views, with each question carefully reviewed for balance and clarity.
+This process reduces the risk of bias in responses, though the brief format of phone surveys limits depth.
 
@@
 
-# Appendix 3 - checks and summary stats {#sec-checks}
+# Appendix 3 Posterior Predictive Check {#sec-checks}
 

### 2024-11-04 paper/paper.qmd
merged commits

--- 2024-11-04 paper/paper.qmd
@@
 We also make other graphs in the Data section, which helps us understand the values themselves.
-
 From the result section we find that the average predicted support for Kamala Harris for the 7 swing states were as follows: Arizona at 46.26%, Nevada at 46.43%, Georgia at 46.59%, North Carolina at 47.26%, Michigan at 47.30%, Pennsylvania at 48.02%, Wisconsin at 48.38%.
@@
 It also helps keep people informed and engaged in the political process, making them more likely to vote and participate in democracy.
-Results paragraph
 
@@
 
-In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin \[CITE NYT PAPER\].
+In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin [CITE NYT PAPER].
 As shown in @fig-supportstate, which displays the support percentages for both Harris and Donald Trump side-by-side by state, the data reveals a competitive race, with relatively close support levels for both candidates in most states.
@@
 ```
-
 @fig-supportAM further examines Harris's support trends in Arizona and Michigan.
@@
 ```
+
 The outcome variable in this model is the proportion of respondents who support Harris in each poll, represented as $y_i$.
@@
 Each coefficient $\beta$ follows the normal distribution with a mean of 0 and a standard deviation of 2.5, providing some regularization while allowing for a range of possible effects.
-The standard deviation $\sigma$ is given an Exponential(1) prior, which constrains it to positive values and provides a realistic expectation of variability around the mean Each predictor captures a unique dimension of voter sentiment:
+The standard deviation $\sigma$ is given an Exponential(1) prior, which constrains it to positive values and provides a realistic expectation of variability around the mean
+Each predictor captures a unique dimension of voter sentiment:
 
@@
 ```
-
 In @fig-modelNPW, we see states of Nevada, Pennsylvania and Wisconsin have a singular declining trend.
@@
 ```
-
-```{r}
-
-```
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-# Results
 
@@
 
-Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.\@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
+Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
 
@@
 
-# Appendix 1
+# Appendix 1 - Pollster Methodology
 
-Why we choose NYT?
-- `numeric_grade` is 3 - `pollscore` is -1.5, a score of reliability called "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.", where negative numbers are better - `transparency_score` is 9, reflects pollsters transparency about their methodology (calculated based on how much information it discloses about its polls and weighted by recency) - `population_full` is 'rv', respondents are registered voters
+The New York Times/Siena College Poll is a respected polling partnership that provides data-driven insights into U.S.
+public opinion, particularly on political, social, and economic issues.
+Known for its rigorous methodology, this poll focuses on capturing the views of registered and likely voters across the country, with particular emphasis on key battleground states during election cycles.
+The poll is frequently cited for its accuracy and transparency, and is widely regarded as reliable by the public, political analysts, and the media.
 
-# Appendix 2
-Here we will be talking about how we would conduct a $100,000 surevey to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent.
+The New York Times/Siena College Poll is conducts live telephone interviews with a national sample of registered voters, using both landlines and cellphones, with over 90% of contacts made via cellphones.
+The sample frame is drawn from a comprehensive voter file covering 200 million registered voters.
+This file provides demographic and geographic information, enabling representative sampling across age, race, and political affiliation.
+The poll targets registered U.S.
+voters, with special focus on battleground states like Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin, determined as swing states.
+Respondents are randomly selected from the voter file, ensuring a diverse sample by region, party, and demographics.
+To mitigate non-response bias, the poll applies demographic weighting, increasing the weight of underrepresented groups, such as young adults or non-college graduates.
+Low response rates (around 2%) present challenges, to mitigate this, weighing adjustments are applied to correct for demographic imbalances and potential non-response bias.
+Questions are structured to be clear, neutral, and representative across political views, with each question carefully reviewed for balance and clarity.
+This process reduces the risk of bias in responses, though the brief format of phone surveys limits depth.
+
+# Appendix 2 - Survey
+
+Here we will be talking about how we would conduct a $100,000 surevy to gather data about this upcoming election to ensure there is minimal error in the data we are collecting and, our data represents the facts we want them to represent.
 Out of the \$100,000 we will allocate \$15000 for the necessary development and administration of the data.
@@
 
-We will not be using telephone surveys because according to reserach [@idSurvey] we find the most people, don't pick up calls from unknown numbers and even the people that pick it up they are less inclined to answer the questions of the survey.
+We will not be using telephone surveys because according to research [@idSurvey] we find the most people, don't pick up calls from unknown numbers and even the people that pick it up they are less inclined to answer the questions of the survey.
 
@@
 Then we take a union of all the samples and we gather a group that we are able to build models based off of.
+
 There are obvious disadvantages as well.
@@
 
-# Appendix 3 - checks and summary stats {#sec-checks}
+# Appendix 3- Posterior Predictive Check {#sec-checks}
 

### 2024-11-04 paper/paper.qmd
fixed refrences

@@
 library(lubridate)
-library(broom)
 library(modelsummary)
@@
 library(knitr)
+library(bayesplot)
 ```
@@
 
-¿ does this go into data cleaning ?
-
-We use the statistical programming language R [@citeR]. Our data comes from [@fivethirtyeight2024] and was cleaned, modeled and graphed, using @broom, @tidy, @rstanarm, @janitor, @lubr, @modelsummary.
+We use the statistical programming language R [@citeR] to compile this paper.
+Our data comes from [@fivethirtyeight2024] and was cleaned and visualized using the packages Tidyverse [@tidy], Lubridate [@lubr], Janitor [@janitor], and Knitr [@knitr] and modeled using Rstanarm [@rstanarm], ModelSummary [@modelsummary] and BayesPlot [@bayesplot].
 
@@
 
-In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin [CITE NYT PAPER].
+In this analysis, we examine polling trends for Kamala Harris across key swing states, focusing on recent polling data from Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin [@NYTswingatates].
 As shown in @fig-supportstate, which displays the support percentages for both Harris and Donald Trump side-by-side by state, the data reveals a competitive race, with relatively close support levels for both candidates in most states.
@@
 Nevertheless, Harris is seen to have had a gradual decline in the state stabilizing at around 46%, which which seems low in comparison to other states that are looking more favorably at Harris than Trump.
-Still, when we look at @supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the democrats.
+Still, when we look at @fig-supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the democrats.
 
@@
 Questions are structured to be clear, neutral, and representative across political views, with each question carefully reviewed for balance and clarity.
-This process reduces the risk of bias in responses, though the brief format of phone surveys limits depth.
+This process reduces the risk of bias in responses, though the brief format of phone surveys limits depth (@NYTpolls).
 
@@
 
-```{r}
+```{r, fig.width=5, fig.height=3}
 #| echo: false
@@
 
-In @fig-ppcheck, we check the Posterior predictive check of model of our model. Since the raw data aligns closely with our prdictions, our model is a good predicton of the data.
+In @fig-ppcheck, we check the Posterior predictive check of model of our model. Since the raw data aligns closely with our predictions, our model is a good prediction of the data.
 

### 2024-11-04 paper/paper.qmd
Discussion

@@
 
+North Carolina has been an exclusively Republican state since 1980, with a singular exception in 2008 when the state was democratic and voted for Obama.
+This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the republican with 49.8% of the vote, with 46.2% going to the Democrats.
+While in 2020, the Republicans maintained their support from the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8%.
+
+Now, if we look at @fig-supportGNN Harris had a spike in early September like many other states where she was polling around 48%, but afterwards had a minor but gradual decline 
+since. Although when we take a look at our model results in @
+
 ## Michigan {#sec-michigan}

### 2024-11-04 paper/paper.qmd
fix

@@
   group_by(end_date_num) %>%
-  summarize(
-    mean_support = mean(pct, na.rm = TRUE),
+  summarize(mean_support = mean(pct, na.rm = TRUE),
     median_support = median(pct, na.rm = TRUE),
     lower_quartile = quantile(pct, 0.25, na.rm = TRUE),
-    upper_quartile = quantile(pct, 0.75, na.rm = TRUE)
-  )
+    upper_quartile = quantile(pct, 0.75, na.rm = TRUE))
 
@@
 
-In states like Arizona, Georgia, Michigan, and North Carolina, the model predicts a steady increase in support for Harris over time, as shown in @fig-modelGAWM.
+In states like Arizona, Georgia, Michigan, and North Carolina, the model predicts a steady increase in support for Harris over time, as shown in @fig-modelGANM.
 In states such as Nevada, Pennsylvania, and Wisconsin, the model predicts a decline in support over time, as seen in @fig-modelNPW.
@@
 #| message: false
-#| label: fig-modelGAWM
+#| label: fig-modelGANM
 #| fig-cap: "Poll predictions of Arizona, Geogria, Michigan, North Carolina"

### 2024-11-04 paper/paper.qmd
discussion North Carolina

@@
 
-harrisdata_clean <- harrisdata %>% drop_na()
-
 harrisdata <- harrisdata %>% drop_na()
@@
 #| message: false
+#| label: tbl-supporttable
+#| tbl-cap: "Averaged predicted support for Harris by state"
 
 expected_probs <- posterior_epred(bay_model, newdata = harrisdata)
-
-# Calculate the median predicted probability for each observation
 median_probs <- apply(expected_probs, 2, median)
-
-# Add median predicted probabilities to `harrisdata`
 harrisdata <- harrisdata %>%
@@
 
-# Calculate the average predicted support by state
 predicted_support_table <- harrisdata %>%
@@
     arrange(average_predicted_support)
-
-#predicted_support_table <- predicted_support_table[, c("state", "average_predicted_support")]
 
@@
 Now, if we look at @fig-supportGNN Harris had a spike in early September like many other states where she was polling around 48%, but afterwards had a minor but gradual decline 
-since. Although when we take a look at our model results in @
+since. Although when we take a look at our model results in @fig-modelGANM we see that our prediction is that Harris's support is predicted to increase in North Carolina.
+
+North Carolina is a significantly difficult state to predict when we consider its Republican history, predicted support for Harris, and @fig-supportstate, where both candidates are neck and neck the chance for this state to go to either candidate is very up in the air.
+
+We believe North Carolina will vote: **Donald Trump**
 

### 2024-11-04 paper/paper.qmd
.

@@
 
-North Carolina has been an exclusively Republican state since 1980, with a singular exception in 2008 when the state was democratic and voted for Obama.
-This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the republican with 49.8% of the vote, with 46.2% going to the Democrats.
-While in 2020, the Republicans maintained their support from the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8%.
-
-Now, if we look at @fig-supportGNN Harris had a spike in early September like many other states where she was polling around 48%, but afterwards had a minor but gradual decline 
-since. Although when we take a look at our model results in @fig-modelGANM we see that our prediction is that Harris's support is predicted to increase in North Carolina.
-
-North Carolina is a significantly difficult state to predict when we consider its Republican history, predicted support for Harris, and @fig-supportstate, where both candidates are neck and neck the chance for this state to go to either candidate is very up in the air.
-
-We believe North Carolina will vote: **Donald Trump**
-
 ## Michigan {#sec-michigan}

### 2024-11-04 paper/paper.qmd
llm added

--- 2024-11-04 paper/paper.qmd
@@
 Now, if we look at @fig-supportGNN Harris had a spike in early September like many other states where she was polling around 48%, but afterwards had a minor but gradual decline 
-since. Although when we take a look at our model results in @fig-modelGANM we see that our prediction is that Harris's support is predicted to increase in North Carolina.
-
-North Carolina is a significantly difficult state to predict when we consider its Republican history, predicted support for Harris, and @fig-supportstate, where both candidates are neck and neck the chance for this state to go to either candidate is very up in the air.
-
-We believe North Carolina will vote: **Donald Trump**
+since. Although when we take a look at our model results in @
 

### 2024-11-04 paper/paper.qmd
North Carolina Edit
Discussion Wisconsin

--- 2024-11-04 paper/paper.qmd
@@
 
+The last swing state we will be discussing is Wisconsinng which we can see in our prediction table, @fig-supporttable, has the high support at 48.38% for Harris from all this swing states we looked at.
+when we look at @fig-supportWP, se a small spike in early september for Harris, followed by a gradual stabilization at around 48% support for Harris similar to what our model shows.
+Now if we also take into ac count @fig-supportstate, we see that Wisconsin has the largest difference in support for Harris in comparison to all previously mentioned swing states.
+
+Lastly, if we look at Wisconsin's voting history, the only time Wisconsin voted Republican since 1988 was in 2016 when the state gave its electoral votes for Trump with 47.2% of the vote to Clinton's 46.5.
+then in 2020 the state went back to blue with Biden winning the state, allt hings considered out of all states we took into account Harris is likely to win.
+
+We believe Wisconsin will vote: **Kamala Harris**
+
+## Tally
+
 ## Weaknesses and next steps

### 2024-11-04 paper/paper.qmd
weakness and next steps

--- 2024-11-04 paper/paper.qmd
@@
 
-## Weaknesses and next steps
+## Limitations and Weaknesses
+- nuemric_grade couldve been used to weigh pollster results, for accuracy 
+- lack of real time adjustments 
+- less data to use becuase we are only considering data since Harris's declaration as a candidate 
+- limited scope of variables (demographic changes, economic conditions, and recent events, are not included)
+- some ambiguity of results, nothing to compare to
 
-Weaknesses and next steps should also be included.
+## Next steps
+
+- use historical data from past elections to identify strong voting patterns 
+- generate confidence intervals 
+- integrate more context, and more vairables : Future studies could examine how state-level socioeconomic and political factors, such as unemployment rates, healthcare concerns, or local election issues, might influence voter preferences in each swing state.
+This would add depth to the model and align predictions more closely with real-world factors.
 

### 2024-11-04 paper/paper.qmd
appendix update

@@
 
-# Appendix 1 - Pollster Methodology
+# Pollster Methodology
 
-The New York Times/Siena College Poll is a respected polling partnership that provides data-driven insights into U.S.
-public opinion, particularly on political, social, and economic issues.
-Known for its rigorous methodology, this poll focuses on capturing the views of registered and likely voters across the country, with particular emphasis on key battleground states during election cycles.
-The poll is frequently cited for its accuracy and transparency, and is widely regarded as reliable by the public, political analysts, and the media.
+The New York Times/Siena College Poll is a respected polling partnership known for providing data-driven insights into U.S.
+public opinion on political, social, and economic issues.
+Recognized for its rigorous methodology, the poll focuses on capturing the views of registered and likely voters across the country, with particular emphasis on key battleground states during election cycles.
+Frequently cited for its accuracy and transparency, it is widely regarded as a reliable source by the public, political analysts, and the media.
 
-The New York Times/Siena College Poll is conducts live telephone interviews with a national sample of registered voters, using both landlines and cellphones, with over 90% of contacts made via cellphones.
-The sample frame is drawn from a comprehensive voter file covering 200 million registered voters.
-This file provides demographic and geographic information, enabling representative sampling across age, race, and political affiliation.
-The poll targets registered U.S.
-voters, with special focus on battleground states like Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin, determined as swing states.
-Respondents are randomly selected from the voter file, ensuring a diverse sample by region, party, and demographics.
-To mitigate non-response bias, the poll applies demographic weighting, increasing the weight of underrepresented groups, such as young adults or non-college graduates.
-Low response rates (around 2%) present challenges, to mitigate this, weighing adjustments are applied to correct for demographic imbalances and potential non-response bias.
-Questions are structured to be clear, neutral, and representative across political views, with each question carefully reviewed for balance and clarity.
-This process reduces the risk of bias in responses, though the brief format of phone surveys limits depth (@NYTpolls).
+The poll conducts live telephone interviews with a national sample of registered voters, reaching over 90% of respondents via cellphones.
+The target population consists of U.S.
+registered voters, with the sampling frame drawn from a comprehensive voter file covering 200 million individuals.
+This file contains demographic and geographic information to ensure sampling is representative across age, race, political affiliation, and region.
+The poll places particular focus on swing states like Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin, where small shifts in voter opinion can determine the electoral outcome.
+Respondents are recruited randomly from the voter file to capture a diverse range of political perspectives, demographics, and regional backgrounds.
+To mitigate non-response bias, which is significant given a response rate of around 2%, the poll applies demographic weighting to underrepresented groups, such as young adults and non-college graduates (@NYTpolls).
 
-# Appendix 2 - Survey
+The methodology of telephone polling provides benefits, such as enabling random sampling across a broad population.
+However, low response rates and the shift away from landlines to cellphones mean certain groups, particularly younger adults, may be underrepresented despite demographic weighting adjustments.
+While the questionnaire is crafted to be clear, neutral, and balanced across political views, the brief format of phone surveys—designed to improve response rates—limits depth.
+Additionally, phone surveys may not reach segments of the population more responsive to digital surveys or in-person interviews, potentially missing out on specific voter perspectives.
+Each respondent’s answer is then adjusted for weighting to ensure proportionality, and results are aggregated to reflect a balanced, representative outcome of public opinion, rather than raw, unadjusted numbers.
+
+Overall, while the New York Times/Siena College Poll is highly regarded for its thorough approach, methodological constraints, such as low response rates and limited outreach to digitally-focused populations, introduce challenges to representativeness.
+
+# Survey
 
@@
 
-# Appendix 3- Posterior Predictive Check {#sec-checks}
+# Posterior Predictive Check {#sec-checks}
 

### 2024-11-04 paper/paper.qmd
updates

@@
 Each poll has two entries, one giving the Data for the polling opinions of Donald Trump and the second for Kamala Harris.
-We have acquired our polling data from @fivethirtyeight2024
+We have acquired our polling data from @fivethirtyeight2024.
 
@@
 
-n predictive modeling, accounting for pollster variability and methodology is essential to avoid overfitting to any single source.
+In predictive modeling, accounting for pollster variability and methodology is essential to avoid overfitting to any single source.
 By including pollster as a predictor in the model, we allow it to recognize that poll results may vary not only due to real changes in public opinion but also because of differences in survey methodologies.
-This consideration enables the model to generate more accurate, generalized predictions, offering a reliable forecast that reflects a well-rounded view of public opinion
+This consideration enables the model to generate more accurate, generalized predictions, offering a reliable forecast that reflects a well-rounded view of public opinion.
 
@@
 
+The Bayesian model we employ leverages probabilistic reasoning to predict Kamala Harris's support across different states.
+Bayesian models offer a flexible framework for incorporating prior information alongside observed data to produce posterior distributions, which reflect updated opinions based on the data.
+This model was chosen for its ability to incorporate prior knowledge, handle uncertainty, and generate predictions that can adapt to the polling data, providing a more comprehensive forecast of Harris’s support across swing states.
+
 We used the `stan_glmer` function from the `rstanarm` package to create a Bayesian regression model with a binomial family and logit link.
@@
 
+The last swing state we will be discussing is Wisconsinng which we can see in our prediction table, @tbl-supporttable, has the high support at 48.38% for Harris from all this swing states we looked at.
+when we look at @fig-supportWP, se a small spike in early september for Harris, followed by a gradual stabilization at around 48% support for Harris similar to what our model shows.
+Now if we also take into ac count @fig-supportstate, we see that Wisconsin has the largest difference in support for Harris in comparison to all previously mentioned swing states.
+
+Lastly, if we look at Wisconsin's voting history, the only time Wisconsin voted Republican since 1988 was in 2016 when the state gave its electoral votes for Trump with 47.2% of the vote to Clinton's 46.5.
+then in 2020 the state went back to blue with Biden winning the state, allt hings considered out of all states we took into account Harris is likely to win.
+
+We believe Wisconsin will vote: **Kamala Harris**
+
+## Electoral Vote Projection and Predicted Outcome
+
+Without considering the swing states and if we assume all the other states will vote as expected Kamala Harris has a slight lead of 226 electoral votes to Donald Trump's 219.
+Now, if we take the 43 combined votes of Arizona, Georgia, and North Carolina, and add it to Trumps tally.
+While allocate Nevada, Michigan, Pennsylvania, and Wisconsin's votes to Harris which sum up to 50 electoral votes.
+This totals to a Democratic victory for Kamala Harris at 276 votes to Donald Trump's 262, we have come to the conclusion that Kamala Harris is likely to win the upcomming United States election by 14 electoral votes.
+
 ## Limitations and Weaknesses
@@
 
-Although there are both advatages and disadvantages, we feel that this survey is well though out and doesn't allow for ambiguity in the answers of the participants.
+Although there are both advantages and disadvantages, we feel that this survey is well though out and doesn't allow for ambiguity in the answers of the participants.
 

### 2024-11-04 paper/paper.qmd
Spelling edit

--- 2024-11-04 paper/paper.qmd
@@
 Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%.
-While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992.
-Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.
+While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant Republican victory in 2016 and a rather marginal win in 2020, which was the first Democratic victory for the state since 1992.
+Georgia is in an interesting situation as lot of the Democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.
 
@@
 
-Nevada is a state who's majority is dessert and wilderness, with two large cities being Reno and Las Vegas, as the case is with many states a Majority of democratic votes come from large cities and the majority of republican votes coming from rural areas.
-Nevada has been a democratic state for the last 4 elections with at least a 2% difference in favor of the democrats in the last two elections.
+Nevada is a state who's majority is dessert and wilderness, with two large cities being Reno and Las Vegas, as the case is with many states a Majority of Democratic votes come from large cities and the majority of Republican votes coming from rural areas.
+Nevada has been a Democratic state for the last 4 elections with at least a 2% difference in favor of the Democrats in the last two elections.
 
-Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the democratic candidates.
+Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the Democratic candidates.
 Nevertheless, Harris is seen to have had a gradual decline in the state stabilizing at around 46%, which which seems low in comparison to other states that are looking more favorably at Harris than Trump.
-Still, when we look at @fig-supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the democrats.
+Still, when we look at @fig-supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the Democrats.
 
@@
 
-North Carolina has been an exclusively Republican state since 1980, with a singular exception in 2008 when the state was democratic and voted for Obama.
-This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the republican with 49.8% of the vote, with 46.2% going to the Democrats.
+North Carolina has been an exclusively Republican state since 1980, with a singular exception in 2008 when the state was Democratic and voted for Obama.
+This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the Republican with 49.8% of the vote, with 46.2% going to the Democrats.
 While in 2020, the Republicans maintained their support from the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8%.
@@
 When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner.
-Although unlike the previously mentioned states 2016 was the first time the state voted for the republican candidate since 1992.
-Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state.
+Although unlike the previously mentioned states 2016 was the first time the state voted for the Republican candidate since 1992.
+Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote Democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state.
 
@@
 Second, Similarly to Nevada and Michigan Pennsylvania Voted for Trump in 2016 with a 48.2% to a 47.5% victory, and for Biden in 2020 with a slightly more significant but not overwhelming victory at 50% to 48.8%.
-2016 was the first time Pennsylvania voted for a republican candidate since 1988.
+2016 was the first time Pennsylvania voted for a Republican candidate since 1988.
 
-Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
+Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more Democratic than Republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
 

### 2024-11-04 paper/paper.qmd
Added image
captions

@@
 #| label: fig-supportstate
-#| fig-cap: "Distribution of Support Percentage by State"
+#| fig-cap: "Distribution of Support Percentage by State for Harris vs. Trump"
 
@@
 #| label: fig-supportWP
-#| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate for Wisconsin and Pennsylvania"
+#| fig-cap:  "Predicted Support Percentage for Harris Over Time in Pennsylvania and Wisconsin Since Candidacy Declaration"
 
@@
 #| label: fig-supportAM
-#| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate"
+#| fig-cap: "Predicted Support Percentage for Harris Over Time in Arizona and Michigan Since Candidacy Declaration"
 
@@
 #| label: fig-supportGNN
-#| fig-cap:  "Support Percentage for Harris by State since Harris declared as Candidate"
+#| fig-cap:  "Predicted Support Percentage for Harris Over Time in Georiga, Nevada, and North Carolina Since Candidacy Declaration"
 
@@
 
-```{r, fig.width= 6, fig.height= 2}
+```{r, fig.width= 5.5, fig.height= 2.5}
 #| echo: false
@@
 #| label: fig-supportpollsters
-#| fig-cap: "Variation in Harris's Support Percentage Across Pollsters Over Time"
+#| fig-cap: "Variation in Harris's Support Percentage Across Pollsters Since Harris' Candidacy Declaration "
 
@@
 #| label: fig-modelGANM
-#| fig-cap: "Poll predictions of Arizona, Geogria, Michigan, North Carolina"
+#| fig-cap: "Predicted Probability of Support for Harris Over Time in Arizona, Georgia, Michigan, and North Carolina Since Harris' Candidacy Declaration"
 
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme(axis.text.x = element_text(hjust = 1)) +
   facet_wrap(~ state) +
@@
 This totals to a Democratic victory for Kamala Harris at 276 votes to Donald Trump's 262, we have come to the conclusion that Kamala Harris is likely to win the upcomming United States election by 14 electoral votes.
+Below we have added a makeup of what we believe the election to look like in an image @270towin:
+
+::: {.cell}
+![Predicted Electoral Votes for the 2024 US Election ](../other/images/Election_map.png){fig-align="center" width="50%"}
+:::
 

### 2024-11-04 paper/paper.qmd
Image adjustment
.

@@
 ::: {.cell}
-![Predicted Electoral Votes for the 2024 US Election ](../other/images/Election_map.png){fig-align="center" width="50%"}
+![Predicted Electoral Votes for the 2024 US Election ](../other/images/Election_map.png){fig-align="center" width="100%"}
 :::
@@
 ## Limitations and Weaknesses
+
+There are several limitations and weaknesses thatcould have affected out research and statistical analysis. First, the variable ****
 - nuemric_grade couldve been used to weigh pollster results, for accuracy 

### 2024-11-04 paper/paper.qmd
more caption changes

@@
 #| label: fig-modelNPW
-#| fig-cap: "Poll predictions of Nevada, Pennsylvania, Wisconsin"
+#| fig-cap: "Predicted Probability of Support for Harris Over Time in Nevada, Pennsylvania, and Wisconsin Since Harris' Candidacy Declaration"
 
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme(axis.text.x = element_text(hjust = 1)) +
   facet_wrap(~ state, ncol = 2) +
@@
 #| label: tbl-supporttable
-#| tbl-cap: "Averaged predicted support for Harris by state"
+#| tbl-cap: "Average Predicted Support for Harris by Swing State"
 
@@
 Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%.
-While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant Republican victory in 2016 and a rather marginal win in 2020, which was the first Democratic victory for the state since 1992.
-Georgia is in an interesting situation as lot of the Democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.
+While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992.
+Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.
 
@@
 
-Nevada is a state who's majority is dessert and wilderness, with two large cities being Reno and Las Vegas, as the case is with many states a Majority of Democratic votes come from large cities and the majority of Republican votes coming from rural areas.
-Nevada has been a Democratic state for the last 4 elections with at least a 2% difference in favor of the Democrats in the last two elections.
+Nevada is a state who's majority is dessert and wilderness, with two large cities being Reno and Las Vegas, as the case is with many states a Majority of democratic votes come from large cities and the majority of republican votes coming from rural areas.
+Nevada has been a democratic state for the last 4 elections with at least a 2% difference in favor of the democrats in the last two elections.
 
-Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the Democratic candidates.
+Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the democratic candidates.
 Nevertheless, Harris is seen to have had a gradual decline in the state stabilizing at around 46%, which which seems low in comparison to other states that are looking more favorably at Harris than Trump.
-Still, when we look at @fig-supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the Democrats.
+Still, when we look at @fig-supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the democrats.
 
@@
 
-North Carolina has been an exclusively Republican state since 1980, with a singular exception in 2008 when the state was Democratic and voted for Obama.
-This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the Republican with 49.8% of the vote, with 46.2% going to the Democrats.
+North Carolina has been an exclusively Republican state since 1980, with a singular exception in 2008 when the state was democratic and voted for Obama.
+This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the republican with 49.8% of the vote, with 46.2% going to the Democrats.
 While in 2020, the Republicans maintained their support from the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8%.
@@
 When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner.
-Although unlike the previously mentioned states 2016 was the first time the state voted for the Republican candidate since 1992.
-Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote Democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state.
+Although unlike the previously mentioned states 2016 was the first time the state voted for the republican candidate since 1992.
+Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state.
 
@@
 Second, Similarly to Nevada and Michigan Pennsylvania Voted for Trump in 2016 with a 48.2% to a 47.5% victory, and for Biden in 2020 with a slightly more significant but not overwhelming victory at 50% to 48.8%.
-2016 was the first time Pennsylvania voted for a Republican candidate since 1988.
+2016 was the first time Pennsylvania voted for a republican candidate since 1988.
 
-Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more Democratic than Republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
+Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
 
@@
 This totals to a Democratic victory for Kamala Harris at 276 votes to Donald Trump's 262, we have come to the conclusion that Kamala Harris is likely to win the upcomming United States election by 14 electoral votes.
-Below we have added a makeup of what we believe the election to look like in an image @270towin:
 
@@
 #| label: fig-ppcheck
-#| fig-cap: "Posterior predictive check of model"
+#| fig-cap: "Posterior Predictive Check of Model"
 

### 2024-11-04 paper/paper.qmd
Edit to weaknesses and limitations

--- 2024-11-04 paper/paper.qmd
@@
 
-There are several limitations and weaknesses thatcould have affected out research and statistical analysis. First, the variable ****
-- nuemric_grade couldve been used to weigh pollster results, for accuracy 
-- lack of real time adjustments 
-- less data to use becuase we are only considering data since Harris's declaration as a candidate 
-- limited scope of variables (demographic changes, economic conditions, and recent events, are not included)
-- some ambiguity of results, nothing to compare to
+There are several limitations and weaknesses that could have affected out research and statistical analysis.
+First, the variable `numeric_grade` which is the variable that weighs pollster results for accuracy could have been used to determing the how we look at each pollster to make sure the pollster we chose was the most accurate of the most available, and could have included it in our model to increase accuracy of our predictions.Second, as the data from polls change often especially leading up to the election it is difficult to make realtime adjustments as new data becomes available.
+Next, Harris only announced her candidacy at the end of July, we did not have a very large time frame to take into account whenn looking at her polling data, as often presidential candidates begin campaigning much earlier than just a few months before the election.
+Further, As we only created a Model for Harris and not Trump we had less to compare to, while we believe our results would not have changed significantly it would have ben halpful to consider a model based around Donald Trump.
+Lastly, in recent history elections have been rather close whether they were close in electoral votes or in the popular vote, pollster are hesitant to state that one candidate will win over another even if their research and data shows otherwise as they are afraid to lose their credibility for future elections if they turn out to be wrong or if their analysis is different than the actual results.
 

### 2024-11-04 paper/paper.qmd
References and spelling edits

@@
 
+Here we will discuss how we believe each state wwill vote both base on data we collected, the models we created and the voting history of each of these swing states in past elections.
+
 ## Arizona {#sec-arizona}
@@
 We can see in @fig-supportAM that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls.
-Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled.
+Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled @towin270.
 
@@
 In @fig-supportGNN at the beginning of the polling data the support for Harris was relatively low, while it did start to grow the polling percentage somewhat plateaued at around 47.5% and does not show any signs of growing.
-Furthermore, in @fig-supportstate we can see that Trump has the edge over Harris considering all of these factors there is a very real possibility that Trump does win Georgia at the end of the day.
+Furthermore, in @fig-supportstate we can see that Trump has the edge over Harris considering all of these factors there is a very real possibility that Trump does win Georgia at the end of the day @towin270.
 
@@
 Nevada is a state who's majority is dessert and wilderness, with two large cities being Reno and Las Vegas, as the case is with many states a Majority of democratic votes come from large cities and the majority of republican votes coming from rural areas.
-Nevada has been a democratic state for the last 4 elections with at least a 2% difference in favor of the democrats in the last two elections.
+Nevada has been a democratic state for the last 4 elections with at least a 2% difference in favor of the democrats in the last two elections @towin270.
 
@@
 This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the republican with 49.8% of the vote, with 46.2% going to the Democrats.
-While in 2020, the Republicans maintained their support from the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8%.
+While in 2020, the Republicans maintained their support from the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8% @towin270.
 
@@
 Although unlike the previously mentioned states 2016 was the first time the state voted for the republican candidate since 1992.
-Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state.
+Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state @towin270.
 
@@
 Second, Similarly to Nevada and Michigan Pennsylvania Voted for Trump in 2016 with a 48.2% to a 47.5% victory, and for Biden in 2020 with a slightly more significant but not overwhelming victory at 50% to 48.8%.
-2016 was the first time Pennsylvania voted for a republican candidate since 1988.
+2016 was the first time Pennsylvania voted for a republican candidate since 1988 @towin270.
 
@@
 
-Lastly, if we look at Wisconsin's voting history, the only time Wisconsin voted Republican since 1988 was in 2016 when the state gave its electoral votes for Trump with 47.2% of the vote to Clinton's 46.5.
-then in 2020 the state went back to blue with Biden winning the state, allt hings considered out of all states we took into account Harris is likely to win.
+Lastly, if we look at Wisconsin's voting history, the only time Wisconsin voted Republican since 1988 was in 2016 when the state gave its electoral votes for Trump with 47.2% of the vote to Clinton's 46.5% then in 2020 the state went back to blue with Biden winning the state, all things considered out of all states we took into account Harris is likely to win @towin270.
 
@@
 This totals to a Democratic victory for Kamala Harris at 276 votes to Donald Trump's 262, we have come to the conclusion that Kamala Harris is likely to win the upcomming United States election by 14 electoral votes.
+Below we have included an image of what we believee the electoral map will look after the election November 5th, @towin270:
 

### 2024-11-04 paper/paper.qmd
.

@@
 Next, Harris only announced her candidacy at the end of July, we did not have a very large time frame to take into account whenn looking at her polling data, as often presidential candidates begin campaigning much earlier than just a few months before the election.
-Further, As we only created a Model for Harris and not Trump we had less to compare to, while we believe our results would not have changed significantly it would have ben halpful to consider a model based around Donald Trump.
+Furthermore, As we only created a Model for Harris and not Trump we had less to compare to, while we believe our results would not have changed significantly it would have been helpful to consider a model based around Donald Trump.
+Additionally we could have looked at the states recent voting history for their members of congress, both the senate , and the house of representatives, as well as the current political party holding power in the state, i.e.
+the party that the governer of each state is affiliated with.
 Lastly, in recent history elections have been rather close whether they were close in electoral votes or in the popular vote, pollster are hesitant to state that one candidate will win over another even if their research and data shows otherwise as they are afraid to lose their credibility for future elections if they turn out to be wrong or if their analysis is different than the actual results.

### 2024-11-04 paper/paper.qmd
Next steps and gramar edits

@@
 There are several limitations and weaknesses that could have affected out research and statistical analysis.
-First, the variable `numeric_grade` which is the variable that weighs pollster results for accuracy could have been used to determing the how we look at each pollster to make sure the pollster we chose was the most accurate of the most available, and could have included it in our model to increase accuracy of our predictions.Second, as the data from polls change often especially leading up to the election it is difficult to make realtime adjustments as new data becomes available.
+First, the variable `numeric_grade` which is the variable that weighs pollster results for accuracy could have been used to determining the how we look at each pollster to make sure the pollster we chose was the most accurate of the most available, and could have included it in our model to increase accuracy of our predictions.Second, as the data from polls change often especially leading up to the election it is difficult to make realtime adjustments as new data becomes available.
 Next, Harris only announced her candidacy at the end of July, we did not have a very large time frame to take into account whenn looking at her polling data, as often presidential candidates begin campaigning much earlier than just a few months before the election.
@@
 Additionally we could have looked at the states recent voting history for their members of congress, both the senate , and the house of representatives, as well as the current political party holding power in the state, i.e.
-the party that the governer of each state is affiliated with.
+the party that the governor of each state is affiliated with.
 Lastly, in recent history elections have been rather close whether they were close in electoral votes or in the popular vote, pollster are hesitant to state that one candidate will win over another even if their research and data shows otherwise as they are afraid to lose their credibility for future elections if they turn out to be wrong or if their analysis is different than the actual results.
@@
 
-- use historical data from past elections to identify strong voting patterns 
-- generate confidence intervals 
+As the election is quickly approachin there are not too many things we would be bale to consider for this election, nevertheless there are a few.
+First, While we did consider the historical voting paterns of each state, we have not done a rigorous analysis with that data, which we would be able to consider when analyzing a the topic in the future.
+As, poll results are inherently uncertain and are not absolute indicators of who will win the election, we could use confidence intervals in the future to demonstrate this uncertainty.
+Next, we could could add more variables into our analysis to create a stronger and more accurate model to help us predict the election with a higher degree of assurance.
+LAstly, we could consider intangibles more heavily in the future, including recent events, both local and worldwide, considering the socioeconomic status of voters and the voting patterns of certain ethnic groups and how the population and demographic has changed since the last election, as well as general sentiment towards the parties and their presidential candidates.
+
 - integrate more context, and more vairables : Future studies could examine how state-level socioeconomic and political factors, such as unemployment rates, healthcare concerns, or local election issues, might influence voter preferences in each swing state.

### 2024-11-04 paper/paper.qmd
proofread introduction

@@
 Our analysis will explore demographic shifts, the impact of key issues, and the level of voter engagement across different swing states.
-Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the forces shaping the 2024 presidential election.
-We use Bayesian regression model to create a model to predict the percent of population that will support in the swing states based on pollster, the number of days since Kamala Harris joined the race.
-More on the model and the variables, in the Data and Model sections.
+Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the opinions shaping the 2024 presidential election.
+Using Bayesian regression modeling, we predict the percent of supprt for Kamala Harris in swing states based on pollster and the number of days since she announced her candidacy.
+Details on our data, variables, and model are discussed further in the Data and Model sections.
 
-We noticed in most predictions of the presidential candidate they are usually based on one survey or pollster and they try to gather data from the one poll to predict the the future President.
-However, through looking at the results from different polls we are able to get a better, more accurate prediction.
-We also look at over state polls which allows us to focus on the swing states.
-Swing states are states that change sides every election.
-The 2 major parties that fight for power are the Republicans and the Democrats.
-Some states are always voting for one party regardless of the candidate so, the swing states are the once that tell us more about who might win the election.
-The gap we are trying to cover is predicting the future President of USA by looking at different polls, while being as effecient as possible by looking only at the swing states.
+Traditional election forecasts often rely on single polls or sources, but our approach combines results from multiple polls to provide a more accurate prediction.
+By focusing on swing states, which historically alternate between parties, we capture the areas most likely to influence the election outcome.
+The 2 major parties that are in competition are the Republicans and the Democrats.
+Some states exhibit a consistent pattern of voting for a particular party, regardless of the candidate.
+In contrast, swing states are more variable, providing key insights into who might win the election.
+Our goal is to predict the outcome of the U.S.
+presidential election by synthesizing data from multiple polls, focusing efficiently on the indecisive swing states.
 
-To go a little more in-depth into what we actually did, we have to first discuss our estimand.
-Our estimand, the parameter we aim to study, is the predicted support for Kamala Harris in the swing states (measured in %).
-We do this by creating Baysian model with 3 parameters; state, pollster, and says since Kamala Harris was announced as candidate.
-We have the priors set with a normal distribution centered at 0, with a standard deviation of 2.5.
-We expect that because the swing states are states that don't have a party that they dominantly vote for, the expected support will be around 50%.
-We also make other graphs in the Data section, which helps us understand the values themselves.
-From the result section we find that the average predicted support for Kamala Harris for the 7 swing states were as follows: Arizona at 46.26%, Nevada at 46.43%, Georgia at 46.59%, North Carolina at 47.26%, Michigan at 47.30%, Pennsylvania at 48.02%, Wisconsin at 48.38%.
+To delve further into our approach, we introduce our estimated - the parameter of interest- which is the predicted support for Kamala Harris in the swing states (measured in %).
+We estimate this by creating Baysian model with 3 parameters; state, pollster, and days since Harris’s candidacy announcement.
+We set priors using a normal distribution centered at 0, with a standard deviation of 2.5.
+Given that swing states tend to lack a dominant party preference, the expected support will be approximately 50%.
+We further explore these variables and their meanings other graphs in the (@sec-data).
+From our analysis in (@sec-results) we find that the average predicted support for Kamala Harris for the 7 swing states were as follows: Arizona at 46.26%, Nevada at 46.43%, Georgia at 46.59%, North Carolina at 47.26%, Michigan at 47.30%, Pennsylvania at 48.02%, Wisconsin at 48.38%.
 Based on this we predict that Arizona, and Georgia will vote Donald trump and the rest of the states will vote Kamala Harris.
 
-Predicting who will become president is important because it helps people understand and prepare for future changes in the country’s policies and priorities.
+Accurate election predictions are important as they help individuals, businesses, and governments anticipate potential shifts in policy, the economy, and international relations.
 A new president can shape laws, affect the economy, influence international relationships, and address key issues like healthcare, education, and climate change.
-Accurate predictions allow citizens, businesses, and governments to plan and adapt.
-For example, if polls indicate a particular candidate might win, groups can start thinking about how their plans might align with the candidate's proposed policies.
-It also helps keep people informed and engaged in the political process, making them more likely to vote and participate in democracy.
+Knowing which candidate is likely to win enables groups to consider how their plans might align with anticipated policies, encouraging public engagement and participation in the democratic process.
+It also helps keep people informed and engaged in the political process, making them more likely to vote and participate in democracy in the future.
 
-The paper is structured with the introduction in the beginning followed by the Data section wich takes a look at the methodology of the paper and how the data was gathered.
-We explain the variables we isolated in order for this analysis to be complete.
-Then it is followed by the Model section where we introduce the model and explain the graphs of related models.
-The result sections will cover the model summary and summary table of the predictions for each of the swing states.
-The last section is the discussion section where we talk about what the results means and we give our final conclusion and weaknesses.
+The paper is structured as follows: the Data section (@sec-data) covers our methodology and data collection process, detailing the variables used in the analysis.
+The Model section (@sec-model) introduces our Bayesian approach and includes related visualizations.
+The Results section (@sec-results) provides a model summary and predicted support percentages for each swing state.
+In the Discussion section (@sec-discussion), we interpret our findings, present final predictions, and discuss limitations, weaknesses and next steps.
 We also give a final word about who will win the upcoming election.
@@
 The second appendix talks about a survey we designed with reasons and disadvantage.
-The third appendix just shows the models in more depth.
+The third appendix shows the Posterior Predictive check of our model.
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-## Model Results
+## Model Results {#sec-results}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 

### 2024-11-04 paper/paper.qmd
Grammar edits

@@
 Through a comprehensive statistical approach, we aim to contribute to the understanding of the evolving electoral landscape and provide meaningful insights into the opinions shaping the 2024 presidential election.
-Using Bayesian regression modeling, we predict the percent of supprt for Kamala Harris in swing states based on pollster and the number of days since she announced her candidacy.
+Using Bayesian regression modeling, we predict the percent of support for Kamala Harris in swing states based on pollsters and the number of days since she announced her candidacy.
 Details on our data, variables, and model are discussed further in the Data and Model sections.
@@
 Accurate election predictions are important as they help individuals, businesses, and governments anticipate potential shifts in policy, the economy, and international relations.
-A new president can shape laws, affect the economy, influence international relationships, and address key issues like healthcare, education, and climate change.
+A new president can affect the previously mentioned factors as well as address key issues like healthcare, education, and climate change.
 Knowing which candidate is likely to win enables groups to consider how their plans might align with anticipated policies, encouraging public engagement and participation in the democratic process.

### 2024-11-04 paper/paper.qmd
edits to paragraph structure
.

@@
 While in the 2020 election Biden ended up winning Arizona with 49.4% of the vote and Trump 49.1%, which is a much smaller margin of victory and what makes Arizona so questionable in the 2024 election.
+
 We can see in @fig-supportAM that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls.
@@
 When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner.
+
 Although unlike the previously mentioned states 2016 was the first time the state voted for the republican candidate since 1992.
@@
 LAstly, we could consider intangibles more heavily in the future, including recent events, both local and worldwide, considering the socioeconomic status of voters and the voting patterns of certain ethnic groups and how the population and demographic has changed since the last election, as well as general sentiment towards the parties and their presidential candidates.
-
-- integrate more context, and more vairables : Future studies could examine how state-level socioeconomic and political factors, such as unemployment rates, healthcare concerns, or local election issues, might influence voter preferences in each swing state.
-This would add depth to the model and align predictions more closely with real-world factors.
 

### 2024-11-04 paper/paper.qmd
discussion proffread

--- 2024-11-04 paper/paper.qmd
@@
 We also give a final word about who will win the upcoming election.
-We also have additional appendices, where we talk about the methodology of the polls and how they compare.
+We also have additional appendices (@sec-app), where we talk about the methodology of the polls and how they compare.
 The second appendix talks about a survey we designed with reasons and disadvantage.
@@
 
-CONFIDENCE INTERVAL TABLE ??
-
 ```{r}
@@
 
-Here we will discuss how we believe each state wwill vote both base on data we collected, the models we created and the voting history of each of these swing states in past elections.
+In this section, we will analyze each state individually, considering our collected data, generated models, and the voting history of each state in previous elections to project likely outcomes.
 
@@
 
-Arizona is a very important swing state especially as the state has switches which party they voted for in in the last two elections with Trump winning 49% of the vote and Clinton with 45.5%, which considering Arizona is a swing state is a rather significant6 margin.
-While in the 2020 election Biden ended up winning Arizona with 49.4% of the vote and Trump 49.1%, which is a much smaller margin of victory and what makes Arizona so questionable in the 2024 election.
-We can see in @fig-supportAM that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls.
-Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled @towin270.
+Arizona remains a pivotal swing state, particularly given its recent shifts in party alignment in the past two elections.
+In 2016, Trump won Arizona with 49% of the vote to Clinton's 45.5%, a relatively substantial margin for a swing state.
+However, in 2020, Biden narrowly won with 49.4% to Trump’s 49.1%, highlighting Arizona's competitiveness for the 2024 election, which is what makes Arizona such an important state.
+As seen in @fig-supportAM, Harris's polling numbers were initially around 42%, but have significantly increased in the most recent polls.
+Despite this trend, @fig-supportstate indicates that Trump still has larger support in the state, with Arizona having the largest polling difference in his favor among the swing states (@towin270).
 
-We believe Arizona will vote: **Donald Trump**
+We project Arizona will vote for: **Donald Trump**
 
@@
 
-Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%.
-While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992.
-Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.
+Georgia's last two elections outcomes were very alike Arizona's outcomes.
+In 2016, Trump secured the state with 50.8% of the vote to Clinton's 45.6%.
+However, Biden won the 2020 election with a slim victory of 49.5% to Trump's 49.3%, which emphasizes Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020.This marked the state’s first Democratic victory since 1992.
+Georgia is a compelling state because most of the democratic vote comes from the Black community, although Harris has seen relatively lower support among Black men.
+Given Biden's last narrow win, even a minor shift in voter preferences could significantly impact the 2024 outcome.
 
-In @fig-supportGNN at the beginning of the polling data the support for Harris was relatively low, while it did start to grow the polling percentage somewhat plateaued at around 47.5% and does not show any signs of growing.
-Furthermore, in @fig-supportstate we can see that Trump has the edge over Harris considering all of these factors there is a very real possibility that Trump does win Georgia at the end of the day @towin270.
+As shown in @fig-supportGNN, Harris’s support rose slightly but plateaued around 47.5%  and does not show any signs of increasing.
+Furthermore, @fig-supportstate shows that Trump currently holds a slight advantage, suggesting a likely outcome in his favor (@towin270).
 
-We believe Georgia will vote: **Donald Trump**
+We project Georgia will vote for: **Donald Trump**
 
@@
 
-We believe Nevada will vote: **Kamala Harris**
+We project Nevada will vote for: **Kamala Harris**
 
@@
 
-We believe North Carolina will vote: **Donald Trump**
+We project North Carolina will vote for: **Donald Trump**
 
@@
 
-We believe Michigan will vote for: **Kamala Harris**
+We project Michigan will vote for: **Kamala Harris**
 
@@
 
-We believe Pennsylvania will vote: **Kamala Harris**
+We project Pennsylvania will vote for: **Kamala Harris**
 
@@
 
-We believe Wisconsin will vote: **Kamala Harris**
+We project Wisconsin will vote for: **Kamala Harris**
 
@@
 
-As the election is quickly approachin there are not too many things we would be bale to consider for this election, nevertheless there are a few.
-First, While we did consider the historical voting paterns of each state, we have not done a rigorous analysis with that data, which we would be able to consider when analyzing a the topic in the future.
+As the election is quickly approaching there are not too many things we would be bale to consider for this election, nevertheless there are a few.
+First, While we did consider the historical voting patterns of each state, we have not done a rigorous analysis with that data, which we would be able to consider when analyzing a the topic in the future.
 As, poll results are inherently uncertain and are not absolute indicators of who will win the election, we could use confidence intervals in the future to demonstrate this uncertainty.
 Next, we could could add more variables into our analysis to create a stronger and more accurate model to help us predict the election with a higher degree of assurance.
-LAstly, we could consider intangibles more heavily in the future, including recent events, both local and worldwide, considering the socioeconomic status of voters and the voting patterns of certain ethnic groups and how the population and demographic has changed since the last election, as well as general sentiment towards the parties and their presidential candidates.
-
-- integrate more context, and more vairables : Future studies could examine how state-level socioeconomic and political factors, such as unemployment rates, healthcare concerns, or local election issues, might influence voter preferences in each swing state.
-This would add depth to the model and align predictions more closely with real-world factors.
+Lastly, we could consider intangibles more heavily in the future, including recent events, both local and worldwide, considering the socioeconomic status of voters and the voting patterns of certain ethnic groups and how the population and demographic has changed since the last election, as well as general sentiment towards the parties and their presidential candidates.
 
@@
 
-# Pollster Methodology
+# Appendix {#sec-app}
+
+## Pollster Methodology
 
@@
 
-# Survey
+## Survey
 
@@
 
-# Posterior Predictive Check {#sec-checks}
+## Posterior Predictive Check {#sec-checks}
 

### 2024-11-04 paper/paper.qmd
.

--- 2024-11-04 paper/paper.qmd
@@
 
-Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.
+Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid September which is when they both becan to plateau at their current polling numbers.
 

### 2024-11-04 paper/paper.qmd
Abstract edit

@@
 date-format: long
-abstract: "We analyse the probability of Kamala Harris winning in the 7 swing states of USA using baysian modeling; Arizona, Michigan, Pennsylvania, Nevada, Georgia, Wisconsin, North Carolina.
+abstract: "We analyse the probability of Kamala Harris winning in the 7 swing states of USA using baysian modeling;the states we look at are Arizona, Michigan, Pennsylvania, Nevada, Georgia, Wisconsin and North Carolina.
 USA is a large country with its national economy affecting the global economic conditions, which is why predicting the future President of USA will help understand the future economic condiotion of the world.
@@
 Provided this we predict that Kamal Harris will win the election.
-By creating a baysian model based on \"polls of polls \" where we compare the results from different polls, we are able to make this prediction."
+By creating a baysian model based on 'polls of 'polls'  where we compare the results from different polls, we are able to make this prediction."
 format:

### 2024-11-04 paper/paper.qmd
pdf update

--- 2024-11-04 paper/paper.qmd
@@
 
-Arizona remains a pivotal swing state, particularly given its recent shifts in party alignment in the past two elections.
-In 2016, Trump won Arizona with 49% of the vote to Clinton's 45.5%, a relatively substantial margin for a swing state.
-However, in 2020, Biden narrowly won with 49.4% to Trump’s 49.1%, highlighting Arizona's competitiveness for the 2024 election, which is what makes Arizona such an important state.
-As seen in @fig-supportAM, Harris's polling numbers were initially around 42%, but have significantly increased in the most recent polls.
-Despite this trend, @fig-supportstate indicates that Trump still has larger support in the state, with Arizona having the largest polling difference in his favor among the swing states (@towin270).
+Arizona is a very important swing state especially as the state has switches which party they voted for in in the last two elections with Trump winning 49% of the vote and Clinton with 45.5%, which considering Arizona is a swing state is a rather significant6 margin.
+While in the 2020 election Biden ended up winning Arizona with 49.4% of the vote and Trump 49.1%, which is a much smaller margin of victory and what makes Arizona so questionable in the 2024 election.
+
+We can see in @fig-supportAM that while the polling numbers for Harris started somewhat low around 42%, the have gone up significantly and have continued to trend upwards in the most recent polls.
+Although, in the @fig-supportstate trump still has larger support in the state, and the largest polling difference in Trump's favour out of all the swing states polled (@towin270).
 
@@
 
-Nevada is a state who's majority is dessert and wilderness, with two large cities being Reno and Las Vegas, as the case is with many states a Majority of democratic votes come from large cities and the majority of republican votes coming from rural areas.
+Nevada, primarily composed of desert and rural areas, sees its Democratic support concentrated in urban centers like Reno and Las Vegas compared to rural areas.
+The state has voted Democratic in the last four elections, with a 2% or greater margin in favor of Democratic candidates in the past two.
 Nevada has been a democratic state for the last 4 elections with at least a 2% difference in favor of the democrats in the last two elections @towin270.

### 2024-11-04 paper/paper.qmd
.

@@
 Provided this we predict that Kamal Harris will win the election.
-By creating a baysian model based on 'polls of 'polls'  where we compare the results from different polls, we are able to make this prediction."
+By creating a baysian model based on 'polls of polls'  where we compare the results from different polls, we are able to make this prediction."
 format:

### 2024-11-04 paper/paper.qmd
spelling edit
discussion

@@
 Through baysian modelling we found that North Carolina (47.26% support for Harris), Nevada(46.43% support for Harris), Wisconsin(48.38% support for Harris), Michigan(47.30% support for Harris), Pennsylvania(48.02% support for Harris) will vote Kamala Harris, while the rest of the swing states might vote Donald Trump, meaning that majority of the swing states will vote Kamala Harris.
-Provided this we predict that Kamal Harris will win the election.
+Provided this we predict that Kamala Harris will win the election.
 By creating a baysian model based on 'polls of polls'  where we compare the results from different polls, we are able to make this prediction."
@@
 The state has voted Democratic in the last four elections, with a 2% or greater margin in favor of Democratic candidates in the past two.
-Nevada has been a democratic state for the last 4 elections with at least a 2% difference in favor of the democrats in the last two elections @towin270.
+Nevada has been a democratic state for the last 4 elections with at least a 2% difference in favor of the democrats in the last two elections (@towin270).
 
-Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the democratic candidates.
-Nevertheless, Harris is seen to have had a gradual decline in the state stabilizing at around 46%, which which seems low in comparison to other states that are looking more favorably at Harris than Trump.
-Still, when we look at @fig-supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the democrats.
+As seen in @fig-supportGNN, Harris experienced a spike in support in early September, peaking above 48%.
+Although her support gradually declined, stabilizing around 46%, @fig-supportstate still shows a marginal lead for Harris over Trump.
+Given Nevada’s recent voting history and current polling, it is probable that Nevada will support Harris (@towin270).
 
@@
 North Carolina has been an exclusively Republican state since 1980, with a singular exception in 2008 when the state was democratic and voted for Obama.
-This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the republican with 49.8% of the vote, with 46.2% going to the Democrats.
-While in 2020, the Republicans maintained their support from the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8% @towin270.
+Thus, North Carolina voted for Trump in the last two elections, where in 2016 there was a more significant win for Republicans with 49.8% of the vote, and 46.2% going to the Democrats.
+While in 2020, the Republicans maintained their support for the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8% (@towin270).
 
-Now, if we look at @fig-supportGNN Harris had a spike in early September like many other states where she was polling around 48%, but afterwards had a minor but gradual decline 
-since. Although when we take a look at our model results in @fig-modelGANM we see that our prediction is that Harris's support is predicted to increase in North Carolina.
-
-North Carolina is a significantly difficult state to predict when we consider its Republican history, predicted support for Harris, and @fig-supportstate, where both candidates are neck and neck the chance for this state to go to either candidate is very up in the air.
+Looking at @fig-supportGNN, Harris had a spike in early September for some states where she was polling around 48%, but afterwards had a minor but gradual decline.
+However, our model results in @fig-modelGANM suggest a potential increase in Harris’s support over time.
+North Carolina remains challenging to predict due to its Republican history and the neck-and-neck nature of the race shown in @fig-supportstate, leaving the state’s outcome highly uncertain.
 
@@
 This totals to a Democratic victory for Kamala Harris at 276 votes to Donald Trump's 262, we have come to the conclusion that Kamala Harris is likely to win the upcomming United States election by 14 electoral votes.
-Below we have included an image of what we believee the electoral map will look after the election November 5th, @towin270:
+Below we have included an image of what we believe the electoral map will look after the election November 5th, @towin270:
 

### 2024-11-04 paper/paper.qmd
final push

@@
 
-Harris is currently leading the polls in Michigan with 49% of the vote and Trump with 47%, this is a rather slim margin, but the first in the discussed states where Harris is leading in the polls.
-In @fig-supportAM we can see that there was a significant jump for Harris and a small dip in October followed by a gradual increase leading up to the elections.
-When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner.
-
-Although unlike the previously mentioned states 2016 was the first time the state voted for the republican candidate since 1992.
-Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state @towin270.
+Harris is currently leading the polls in Michigan with 49% of the vote compared to Trump’s 47%.
+This is a slim margin, but it is the first of the states we've discussed where Harris is leading in the polls.
+In @fig-supportAM, we see a significant jump in support for Harris, followed by a small dip in October, and then a gradual increase leading up to the election.
+Looking at the state’s voting history in the last two elections, Michigan has aligned with Arizona and Georgia, voting for Biden in 2020 and for Trump in 2016, with vote percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3% in 2016, respectively.
+Unlike the previously mentioned states, however, 2016 was the first time since 1992 that Michigan voted for a Republican candidate.
+Considering Michigan is home to Detroit, the largest Black-majority city in the United States, and has a large population of union workers who often support the Democratic party due to its pro-union stance, it is likely that Harris will win the state (@towin270).
 
@@
 
-Pennsylvania is an interesting for several reasons, the first is that before Tim Walz was announced as Kamala Harris's running mate, one of the speculated candidates was Josh Shapiro, the current governor of Pennsylvania, which presidential candidates often consider from where their potential running mates are from as they tend to help with swing states nevertheless, she did not choose Shapiro as her running mate and this is no longer a factor.
-
-Second, Similarly to Nevada and Michigan Pennsylvania Voted for Trump in 2016 with a 48.2% to a 47.5% victory, and for Biden in 2020 with a slightly more significant but not overwhelming victory at 50% to 48.8%.
-2016 was the first time Pennsylvania voted for a republican candidate since 1988 @towin270.
-
-Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid September which is when they both becan to plateau at their current polling numbers.
+Pennsylvania is an interesting case for several reasons.
+First, before Tim Walz was announced as Kamala Harris's running mate, one of the speculated candidates was Josh Shapiro, the current governor of Pennsylvania.
+Presidential candidates often consider the home states of potential running mates, as they can help with swing states.
+However, Harris ultimately did not choose Shapiro, so this is no longer a factor.
+Second, similar to Nevada and Michigan, Pennsylvania voted for Trump in 2016 with a narrow margin of 48.2% to 47.5%, and for Biden in 2020 with a slightly larger but still close victory of 50% to 48.8%.
+Notably, 2016 was the first time Pennsylvania voted for a Republican candidate since 1988 (@towin270).
+Lastly, if we look at @fig-supportWP, we can see that Harris has been polling consistently at around 48% since mid-September, which is somewhat similar to the trend in Georgia.
+However, Pennsylvania appears to be leaning more Democratic than Republican compared to Georgia.
+In @fig-supportGNN, both states show a similar increase in support for Harris since polling began, reaching a plateau in mid-September at their current polling numbers.
 
@@
 
-The last swing state we will be discussing is Wisconsinng which we can see in our prediction table, @tbl-supporttable, has the high support at 48.38% for Harris from all this swing states we looked at.
-when we look at @fig-supportWP, se a small spike in early september for Harris, followed by a gradual stabilization at around 48% support for Harris similar to what our model shows.
-Now if we also take into ac count @fig-supportstate, we see that Wisconsin has the largest difference in support for Harris in comparison to all previously mentioned swing states.
+The last swing state we will be discussing is Wisconsin, which, as we can see in our prediction table (@tbl-supporttable), has the highest support for Harris at 48.38% among all the swing states we examined.
+Looking at @fig-supportWP, we see a small spike in early September for Harris, followed by a gradual stabilization at around 48% support, consistent with our model’s projections.
+Additionally, in @fig-supportstate, Wisconsin shows the largest difference in support for Harris compared to all the previously discussed swing states.
 
-Lastly, if we look at Wisconsin's voting history, the only time Wisconsin voted Republican since 1988 was in 2016 when the state gave its electoral votes for Trump with 47.2% of the vote to Clinton's 46.5% then in 2020 the state went back to blue with Biden winning the state, all things considered out of all states we took into account Harris is likely to win @towin270.
+Finally, if we consider Wisconsin's voting history, the only time it voted Republican since 1988 was in 2016, when the state gave its electoral votes to Trump with 47.2% of the vote, compared to Clinton's 46.5%.
+In 2020, the state returned to blue, with Biden winning.
+Considering all factors, Harris is likely to win Wisconsin and reach the 270 electoral votes needed to win the election (@towin270).
 
@@
 
-Without considering the swing states and if we assume all the other states will vote as expected Kamala Harris has a slight lead of 226 electoral votes to Donald Trump's 219.
-Now, if we take the 43 combined votes of Arizona, Georgia, and North Carolina, and add it to Trumps tally.
-While allocate Nevada, Michigan, Pennsylvania, and Wisconsin's votes to Harris which sum up to 50 electoral votes.
-This totals to a Democratic victory for Kamala Harris at 276 votes to Donald Trump's 262, we have come to the conclusion that Kamala Harris is likely to win the upcomming United States election by 14 electoral votes.
-Below we have included an image of what we believe the electoral map will look after the election November 5th, @towin270:
+Without considering the swing states and assuming all other states will vote as expected, Kamala Harris has a slight lead of 226 electoral votes to Donald Trump's 219.
+Now, if we allocate the 43 combined votes of Arizona, Georgia, and North Carolina to Trump and assign the votes of Nevada, Michigan, Pennsylvania, and Wisconsin to Harris, which total 50 electoral votes, the result is a Democratic victory for Kamala Harris with 276 votes to Donald Trump's 262.
+We conclude that Kamala Harris is likely to win the upcoming United States election by 14 electoral votes.
+Below, we have included an image of our projected electoral map for the November 5th election (@towin270):
 
@@
 
-There are several limitations and weaknesses that could have affected out research and statistical analysis.
-First, the variable `numeric_grade` which is the variable that weighs pollster results for accuracy could have been used to determining the how we look at each pollster to make sure the pollster we chose was the most accurate of the most available, and could have included it in our model to increase accuracy of our predictions.Second, as the data from polls change often especially leading up to the election it is difficult to make realtime adjustments as new data becomes available.
-Next, Harris only announced her candidacy at the end of July, we did not have a very large time frame to take into account whenn looking at her polling data, as often presidential candidates begin campaigning much earlier than just a few months before the election.
-Furthermore, As we only created a Model for Harris and not Trump we had less to compare to, while we believe our results would not have changed significantly it would have been helpful to consider a model based around Donald Trump.
-Additionally we could have looked at the states recent voting history for their members of congress, both the senate , and the house of representatives, as well as the current political party holding power in the state, i.e.
-the party that the governor of each state is affiliated with.
-Lastly, in recent history elections have been rather close whether they were close in electoral votes or in the popular vote, pollster are hesitant to state that one candidate will win over another even if their research and data shows otherwise as they are afraid to lose their credibility for future elections if they turn out to be wrong or if their analysis is different than the actual results.
+There are several limitations and weaknesses that could have affected our research and statistical analysis.
+First, the variable `numeric_grade`, which weighs pollster results for accuracy, could have been better utilized to determine how we evaluate each pollster.
+By ensuring that we selected the most accurate pollsters available and incorporating this into our model, we might have increased the accuracy of our predictions.
+
+Second, since polling data often changes, especially leading up to an election, it is challenging to make real-time adjustments as new data becomes available.
+Additionally, because Harris only announced her candidacy at the end of July, we did not have a substantial time frame to analyze her polling data.
+Typically, presidential candidates start campaigning much earlier, which gives a more comprehensive data set for analysis.
+
+Moreover, since we only created a model for Harris and not for Trump, we had limited comparisons.
+While we believe our results would not have significantly changed, having a model based on Trump could have provided a more balanced analysis.
+We also could have benefited from examining recent voting histories in each state for members of Congress, both in the Senate and the House of Representatives, as well as the current political party in power, such as the party affiliation of each state’s governor.
 
@@
 
-As the election is quickly approaching there are not too many things we would be bale to consider for this election, nevertheless there are a few.
-First, While we did consider the historical voting patterns of each state, we have not done a rigorous analysis with that data, which we would be able to consider when analyzing a the topic in the future.
-As, poll results are inherently uncertain and are not absolute indicators of who will win the election, we could use confidence intervals in the future to demonstrate this uncertainty.
-Next, we could could add more variables into our analysis to create a stronger and more accurate model to help us predict the election with a higher degree of assurance.
-Lastly, we could consider intangibles more heavily in the future, including recent events, both local and worldwide, considering the socioeconomic status of voters and the voting patterns of certain ethnic groups and how the population and demographic has changed since the last election, as well as general sentiment towards the parties and their presidential candidates.
+As the election quickly approaches, there are not many additional factors we can consider at this stage.
+However, there are still a few areas for future improvement.
+First, while we did take into account the historical voting patterns of each state, we did not perform a rigorous analysis of that data.
+This is something we could explore more deeply in future studies.
+Additionally, because poll results are inherently uncertain and are not absolute predictors of the election outcome, we could use confidence intervals in the future to better illustrate this uncertainty.
+
+Next, we could incorporate more variables into our analysis to create a stronger and more accurate model, which would help us make more reliable predictions.
+Lastly, we should consider intangibles more thoroughly, such as recent events both locally and globally, the socioeconomic status of voters, and the voting behavior of different ethnic groups.
+It would also be beneficial to analyze how population and demographic shifts since the last election might impact voting trends, as well as general public sentiment toward the political parties and their presidential candidates.
 
